---
title: "Chapter21"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 21. "Iteration" [Chapter 17 hardcopy]
### 21.2.1 Exercises

Q1.
Write 'for' loops to:

* Compute the mean of every column in `mtcars`
* Determine the type of each column in `nycflights13::flights`
* Compute the number of unique values in each column of `iris`
* Generate 10 random normals for each of `μ = −10, 0, 10`, and `100`

Think about the output, sequence, and body before you start writing the loop.

A1.


-----------

Q2.
Eliminate the 'for' loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))

x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
```

A2.


-----------

Q3. Combine your function writing and 'for' loop skills:

* Write a for loop that `prints()` the lyrics to the children’s song “Alice the camel”.
* Convert the nursery rhyme “ten in the bed” to a function. Generalise it to any number of people in any sleeping structure.
* Convert the song “99 bottles of beer on the wall” to a function. Generalise to any number of any vessel containing any liquid on any surface.

A3.


-----------

Q4.
It’s common to see 'for' loops that don’t preallocate the output and instead increase the length of a vector at each step:
```{r}
output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output
```

How does this affect performance? Design and execute an experiment.

A4.


-----------

## 21. "Iteration" [Chapter 17 hardcopy]
### 21.3.5 Exercises


Q1.
Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, `files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now want to read each one with `read_csv()`. Write the 'for' loop that will load them into a single data frame.

A1.


-----------

Q2.

* What happens if you use `for (nm in names(x))` and `x` has no names?
* What if only some of the elements are named?
* What if the names are not unique?

A2.

-----------

Q3. Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, `show_mean(iris)` would print:

```{r}
show_mean(iris)
#> Sepal.Length: 5.84
#> Sepal.Width:  3.06
#> Petal.Length: 3.76
#> Petal.Width:  1.20
```

(Extra challenge: what function did I use to make sure that the numbers lined up nicely, even though the variable names had different lengths?)


A3.

-----------

Q4. What does this code do? How does it work?

```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

A4.


-----------

## 21. "Iteration" [Chapter 17 hardcopy]
### 21.4.1 Exercises

Q1. Read the documentation for `apply()`. In the 2d case, what two 'for' loops does it generalise?

A1.

-----------

Q2. Adapt `col_summary()` so that it only applies to numeric columns You might want to start with an `is_numeric()` function that returns a logical vector that has a `TRUE` corresponding to each numeric column.

A2.


-----------

## 21. "Iteration" [Chapter 17 hardcopy]
### 21.5.3 Exercises

Q1.
[cf. 21.2.1 Exercises] Write code that uses one of the map functions to:

* Compute the mean of every column in `mtcars`
* Determine the type of each column in `nycflights13::flights`
* Compute the number of unique values in each column of `iris`
* Generate 10 random normals for each of `μ = −10, 0, 10`, and `100`

A1.


-----------

Q2. How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

A2.

-----------

Q3.

* What happens when you use the map functions on vectors that aren’t lists?
* What does `map(1:5, runif)` do? Why?

A3.


-----------

Q4.

* What does `map(-2:2, rnorm, n = 5)` do? Why?
* What does `map_dbl(-2:2, rnorm, n = 5)` do? Why?

A4.


-----------

Q5. Rewrite `map(x, function(df) lm(mpg ~ wt, data = df))` to eliminate the anonymous function.

A5.


-----------

## 21. "Iteration" [Chapter 17 hardcopy]
### 21.9.3 Exercises

Q1.
Implement your own version of `every()` using a 'for' loop. Compare it with `purrr::every()`. What does purrr’s version do that your version doesn’t?

A1.

-----------

Q2. Create an enhanced `col_sum()` that applies a summary function to every numeric column in a data frame.

A2.

-----------

Q3.

A possible base R equivalent of `col_sum()` is:
```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```

But it has a number of bugs as illustrated with the following inputs:
```{r}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```


What causes the bugs?

A3.



