---
title: "Chapter27"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 27. "R Markdown" [Chapter 21 hardcopy]
### 27.2.1 Exercises

Q1.
Create a new notebook using `File > New File > R Notebook`. Read the instructions. Practice running the chunks. Verify that you can modify the code, re-run it, and see modified output

A1. <See separate file>
Running code in RStudio updates the output (eg. plot) there. Saving it (Cmd-S) creates `*.nb.html` version in active directory

"The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed." -- Instructions in an R Notebook template

-----------

Q2.
Create a new R Markdown document with `File > New File > R Markdown`… Knit it by clicking the appropriate button. Knit it by using the appropriate keyboard short cut. Verify that you can modify the input and see the output update

A2. <See separate file>
Prompted to give the `*.Rmd` file a title (and choose default output format) when creating it. Saving it doesn't create an html (etc) file. Knitting it to html did create `*.html` version in active directory, and open the file

-----------

Q3.
Compare and contrast the R notebook and R Markdown files you created above.
How are the outputs similar? How are they different?
How are the inputs similar? How are they different?
What happens if you copy the YAML header from one to the other?

A3.
[http://rmarkdown.rstudio.com/r_notebooks.html]

* Outputs appear similar/'expected' (though the default boilerplate code is different)
* Header for notebook was a generic [`title: "R Notebook"`] and [`output: html_notebook`]
* Header for R Markdown file was as selected when creating [`title: "R4DS_Wk13"`] and [`output: html_document`]
* Couldn't be sure of any/what effect when copy&paste YAML header from one to the other (and vice versa). [But what should happen is that `output:` determines what type of document it is]

"R notebook files show the output inside the editor, while hiding the console. R markdown files shows the output inside the console, and does not show output inside the editor" -- [https://jrnold.github.io/r4ds-exercise-solutions/r-markdown.html]

-----------

Q4.
Create one new R Markdown document for each of the three built-in formats: HTML, PDF and Word.
Knit each of the three documents. How does the output differ? How does the input differ?
(You may need to install LaTeX in order to build the PDF output — RStudio will prompt you if this is necessary.)

A4. <See separate files>

* I used same (existing) R Markdown document, but knit to pdf and to Word. The YAML header `output:` changed
* Was prompted to download TeX [`"Mac OS X: TexLive 2013 (Full) - http://tug.org/mactex/
(NOTE: Download with Safari rather than Chrome _strongly_ recommended)"`] ...which I did (3.0GB --> 5.48GB when unpacked)
* After restarting RStudio, could complete the knit to pdf. Knit to Word was ok

"They produce different outputs, both in the final documents and intermediate files (notably the type of plots produced). The only difference in the inputs is the value of output in the YAML header: word_document for Word documents, pdf_document for PDF documents, and html_document for HTML documents" -- [https://jrnold.github.io/r4ds-exercise-solutions/r-markdown.html]


-----------

## 27. "R Markdown" [Chapter 21 hardcopy]
### 27.3.1 Exercises

Q1.
Practice what you’ve learned by creating a brief CV. The title should be your name, and you should include headings for (at least) education or employment. Each of the sections should include a bulleted list of jobs/degrees. Highlight the year in bold

A1. <See separate file>

[`@eringrand : If you're using a RNotebook, once you "preview it" (i.e knit it for the first time) it'll update on its own every time you Save the document`]

-----------

Q2.
Using the R Markdown quick reference, figure out how to:

1. Add a footnote
2. Add a horizontal rule
3. Add a block quote

A2. <ALSO see separate file>

1. eg. Within the text: `[^1]`. At bottom of the document: `[^1]: Here is the footnote`
2. "Horizontal Rule / Page Break : Three or more asterisks or dashes" ...`***` or `---`
3. Blockquote : each line preceded by a `>` character and an optional space

-----------

Q3.
Copy and paste the contents of `diamond-sizes.Rmd` from [https://github.com/hadley/r4ds/tree/master/rmarkdown] in to a local R markdown document. Check that you can run it, then add text after the frequency polygon that describes its most striking features.

A3. <See separate file. NOTE: I changed display of `diamond-sizes.Rmd` to Raw view, to be able to copy the formatting>


-----------

## 27. "R Markdown" [Chapter 21 hardcopy]
### 27.4.7 Exercises
  
Q1.
Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting `echo = FALSE` on each chunk, set a global option

A1.

"echo = FALSE prevents code, but not the results from appearing in the finished file. Use this when writing reports aimed at people who don’t want to see the underlying R code" -- s27.4.2 R4DS book

"As you work more with **knitr**, you will discover that some of the default chunk options don’t fit your needs and you want to change them. You can do this by calling `knitr::opts_chunk$set()` in a code chunk" -- s27.4.5 R4DS book

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)

# "That will hide the code by default, so only showing the chunks you deliberately choose to show (with `echo = TRUE`)" -- s27.4.5 R4DS book
```


-----------

Q2.
Download `diamond-sizes.Rmd` from [https://github.com/hadley/r4ds/tree/master/rmarkdown]. Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes

A2.
```{r}
?kable

diamonds %>%
  select(carat, cut, color, clarity, price) %>%
  arrange(desc(carat)) %>%
  head(20) %>%
  knitr::kable()                                   # re. s27.4.3 R4DS book
```


-----------

Q3.
Modify `diamonds-sizes.Rmd` to use `comma()` to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.

A3.
`comma()` was a 'helper function' created in s27.4.6 R4DS book :
```{r}
comma <- function(x) format(x, digits = 2, big.mark = ",")
```

`smaller` was defined in s27.2 R4DS book :
```{r}
smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

SO THEN : using inline code, (with `r `), like in s27.4.6 R4DS book :

We have data about `r nrow(diamonds)` diamonds. Only `r comma((nrow(diamonds) - nrow(smaller))/nrow(diamonds) * 100)`% are larger than 2.5 carats.

..."When the report is knit, the results of these computations are inserted into the text"

However [https://jrnold.github.io/r4ds-exercise-solutions/r-markdown.html#code-chunks] suggests :
"I moved the computation of the number larger and percent of diamonds larger than 2.5 carats into a code chunk. I find that it is best to keep inline R expressions simple, usually consisting of an object and a formatting function. This makes it both easier to read and test the R code, while simultaneously making the prose easier to read. It helps the readability of the code and document to keep the computation of objects used in prose close to their use. Calculating those objects in a code chunk with the include = FALSE option (as is done in `diamonds-size.Rmd`) is useful in this regard."


-----------

Q4.
Set up a network of chunks where `d` depends on `c` and `b`, and both `b` and `c` depend on `a`. Have each chunk print `lubridate::now()`, set `cache = TRUE`, then verify your understanding of caching

A4.

* `a` has no dependencies
* `b` depends on `a`
* `c` depends on `a`
* `d` depends on `c` and `b`

Based on [https://jrnold.github.io/r4ds-exercise-solutions/r-markdown.html#code-chunks], with `w`, `x`, `y`, `z` there to 'verify your understanding' :
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r a}
print(lubridate::now())
x <- 1
```

```{r b, dependson = c("a")}
print(lubridate::now())
y <- x + 1
```

```{r c, dependson = c("a")}
print(lubridate::now())
z <- x * 2
```

```{r d, dependson = c("c", "b")}
print(lubridate::now())
w <- y + z
```




