---
title: "Chapter16"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
```


## 16. "Dates and times" [Chapter 13 hardcopy]
### 16.2.4 Exercises

Q1.
What happens if you parse a string that contains invalid dates?
```{r}
ymd(c("2010-10-10", "bananas"))
```

A1.
Results in an NA with warning message :
`1 failed to parse.[1] "2010-10-10" NA`

-----------

Q2.
What does the `tzone` argument to `today()` do? Why is it important?

A2.
```{r}
?today
```

"a character vector specifying which time zone you would like to find the current date of. `tzone` defaults to the system time zone set on your computer" eg. `today("GMT")` -- Help

Different time-zones can have different dates, so value of `today()` might vary ...And so it is important to know which time-zone is the 'point of view'

-----------

Q3.
Use the appropriate **lubridate** function to parse each of the following dates:
```{r}
d1 <- "January 1, 2010"                             # Q3.1
d2 <- "2015-Mar-07"                                 # Q3.2
d3 <- "06-Jun-2017"                                 # Q3.3
d4 <- c("August 19 (2015)", "July 1 (2015)")        # Q3.4
d5 <- "12/30/14"       # Dec 30, 2014               # Q3.5
```

A3.
```{r}
mdy(d1)                                             # A3.1
ymd(d2)                                             # A3.2
dmy(d3)                                             # A3.3
mdy(d4)                                             # A3.4
mdy(d5)                                             # A3.5
```


----------

## 16. "Dates and times" [Chapter 13 hardcopy]
### 16.3.4 Exercises
  
Q1.
How does the distribution of flight times within a day change over the course of the year?

A1.
Most obvious way to do this would be to compare daily distributions from during the year (either from a sample of individual days, or weekly/monthly/quarterly aggregates). Is this what the question means? (Else, if we look at how every day changes, I think we'd need to use an average (mean, SD) for the daily distribution, to plot it)

In 16.2.2 (R4DS book) `flights_dt` was defined (and would then used in plots) :
```{r}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))
```

From s16.2.2 (R4DS book) we had 'distribution of departure times within a single day' :
```{r}
flights_dt %>%                            # `flights_dt` shown in code block above, for this
  filter(dep_time < ymd(20130102)) %>%
  ggplot(aes(dep_time)) +
  geom_freqpoly(binwidth = 600)           # 600 seconds = 10 minutes
```

And from s16.3.3 (R4DS book) we had 'distribution of flights across the course of the day for every day of the year' :
```{r}
flights_dt %>%
  mutate(dep_hour = update(dep_time, yday = 1)) %>%  # All 365 days treated as 1/1/13 (`yday=1`)
  ggplot(aes(dep_hour)) +
    geom_freqpoly(binwidth = 300)                    # 300 seconds = 5 minutes
```

From [https://jrnold.github.io/r4ds-exercise-solutions/dates-and-times.html#date-time-components] : "Let’s try plotting this by month" :
```{r}
flights_dt %>%
  mutate(time = hour(dep_time) * 100 + minute(dep_time),  # max value 2359?
         mon = as.factor(month(dep_time))                 # monthly aggregate of 'within a day'
         ) %>%
  ggplot(aes(x = time, color = mon)) +                    # No need for `group = mon` too
  geom_freqpoly(binwidth = 100)                           # 100 minutes
```

...It continues : "This will look better if everything is normalized within groups. The reason that February is lower is that there are fewer days and thus fewer flights". And it plots that

-----------

Q2.
Compare `dep_time`, `sched_dep_time` and `dep_delay`. Are they consistent? Explain your findings

A2.
```{r}
nycflights13::flights %>%
  ggplot(aes(dep_time, sched_dep_time)) +
  geom_point()
```

This plot shows most points along or near-below a diagonal line, indicating on-time departure (or small delay).
The question is probably asking whether `sched_dep_time + dep_delay = dep_time`
(cf. In a previous Chapter we found that `dep_time + air_time != arr_time` ...Delay on tarmac?)
```{r}
glimpse(flights)
```

XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q3.
Compare `air_time` with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport)

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q4.
How does the average delay time change over the course of a day? Should you use `dep_time` or `sched_dep_time`? Why?

A4.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q5.
On what day of the week should you leave if you want to minimise the chance of a delay?

A5.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q6.
What makes the distribution of `diamonds$carat` and `flights$sched_dep_time` similar?

A6.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q7.
Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. (Hint: create a binary variable that tells you whether or not a flight was delayed)

A7.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 16. "Dates and times" [Chapter 13 hardcopy]
### 16.4.5 Exercises

Q1. Why is there `months()` but no `dmonths()`?

A1. There is no standard length of month (30, 31 days), so it wouldn't make sense to define it in seconds

-----------

Q2.
Explain `days(overnight * 1)` to someone who has just started learning R. How does it work?

A2.
This is referring to some code in s16.4.2, designed to add '`days(1)` to the arrival time of each overnight flight'...
```{r}
mutate(
  overnight = arr_time < dep_time,
  arr_time = arr_time + days(overnight * 1),
  sched_arr_time = sched_arr_time + days(overnight * 1)
)
# We only want to add `days(1)` in cases where we omitted it before. `overnight` is TRUE (1) or FALSE (0).
# So if it is an overnight flight, we add 1 day, and if not, no days are added
```

-----------

Q3.
Q3.1 Create a vector of dates giving the first day of every month in 2015.
Q3.2 Create a vector of dates giving the first day of every month in the current year

A3.1
```{r}
ymd(c("2015-01-01", "2015-02-10", "2015-03-01", "2015_04-01", "2015-05-01", "2015-06-01", "2015-07-01", "2015-08-10", "2015-09-01", "2015_10-01", "2015-11-01", "2015-12-01"))

# That worked, but could be done more concisely :
ymd("2015-01-01") + months(0:11)
```


A3.2
Again, could do it the 'long' way or 'concise' way, as above, just replacing 2015 with 2017. BUT that seems trivial.

INSTEAD, assume that the question is asking us to code for a generic 'current year' :
```{r}
# "I can do that by taking `today()` and truncating it to the year using `floor_date`" :
floor_date(today(), unit = "year") + months(0:11)
# -- [https://jrnold.github.io/r4ds-exercise-solutions/dates-and-times.html#time-spans]
```

-----------

Q4.
Write a function that given your birthday (as a date), returns how old you are in years

A4.
We haven't covered Functions yet (s19). This is from [https://jrnold.github.io/r4ds-exercise-solutions/dates-and-times.html#time-spans] :
```{r}
age <- function(bday) {
  (bday %--% today()) %/% years(1)
}
age(ymd("1969-07-29"))
# The parts between `{}` were covered in s16.4.3 R4DS book
```

-----------

Q5.
Why can’t `(today() %--% (today() + years(1)) / months(1)` work?

A5.
```{r}
(today() %--% (today() + years(1)) / months(1)
# Editor shows a red cross, with "unmatched opening bracket `(`" ...but I don't think that was the intended error ...?
```

```{r}
(today() %--% (today() + years(1))) / months(1)
# This works and gives an answer of 12

(today() %--% (today() + years(1)) / months(1))
# This works and gives an answer of 12
```


