---
title: "Chapter12"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 12. "Tidy data" [Chapter 9 hardcopy]
### 12.2.1 Exercises

Q1. Using prose, describe how the variables and observations are organised in each of the sample tables.

A1.
```{r}
?table1
```

s12.2 "Tidy data" :
"Each dataset shows the same values of four variables `country`, `year`, `population`, and `cases`, but each dataset organises the values in a different way" :

* `table1`:     TIDY: Each variable has its own column. Each observation has its own row. (By `country`, by `year`)
* `table2`: NOT TIDY: 2 rows per observation : `cases` & `population` are rows in `type` column, instead of being their own cols
* `table3`: NOT TIDY: `cases` & `population` are combined in `rate` column, instead of having their own cols, (so each value doesn't have its own cell, either)
* `table4`: NOT TIDY: 2 observations per row : `1999` & `2000` are cols, instead of being in a `year` column. (There are actually 2 tables, for `cases` & `population` ...`table4a` & `table4b`)

-----------

Q2.
Compute the `rate` for `table2`, and `table4a` + `table4b`. You will need to perform four operations:

1. Extract the number of TB cases per country per year       # cases
2. Extract the matching population per country per year      # population
3. Divide cases by population, and multiply by 10000         # rate
4. Store back in the appropriate place
Which representation is easiest to work with? Which is hardest? Why?

A2.
Some notes, from previous chapters, about where we met the `filter()` and `$` used below (for TABLE2 and TABLE4) :

* [s5.2 Filter rows with `filter()`] : "`filter()` allows you to subset observations based on their values"
* [s5.4 Select columns with `select()`] : "`select()` allows you to rapidly zoom in on a useful subset"
* [s10.3.2 Subsetting] : "If you want to pull out a single variable, you need some new tools, `$` and `[[`"
* [s10.4 Interacting with older code] : "We don’t use `[` much in this book because `dplyr::filter()` and `dplyr::select()` allow you to solve the same problems with clearer code"


TABLE2 failed attempt - thinking it would be easy :
```{r}
cases <- table2 %>% filter(type == "cases")                  # 1.
population <- table2 %>% filter(type == "population")        # 2.
rate <- cases / population                                   # 3.
# FAIL : "Error in FUN(left, right) : non-numeric argument to binary operator"
```

TABLE2 building a tibble up from pieces. Assumes initial 4 rows for each country, all sorted :
```{r}
cases <- filter(table2, type == "cases")$count               # 1. ...This time, take only the relevant column
population <- filter(table2, type == "population")$count     # 2. ...This time, take only the relevant column
country <- filter(table2, type == "cases")$country           #    ...`filter()` by one `type`, either one
year <- filter(table2, type == "cases")$year                 #    ...`filter()` by one `type`, either one
table2_new <- tibble(country = country,
                        year = year,
                       cases = cases,
                  population = population,
                        rate = cases / population * 10000)   # 3., 4. ...`rate` to answer the Q, rest for future use
table2_new
```


TABLE4 building a tibble up from pieces. Assumes `table4a` & `table4b` are the same length, and sorted in the same order :
```{r}
table4_new <- tibble(country = table4a$country,              #    ...`country` from one table, either one
       `1999` = table4a$`1999` / table4b$`1999` * 10000,     # 3.
       `2000` = table4a$`2000` / table4b$`2000` * 10000)     # 3., 4. ...`rate` answers the Q. But no tidy dataset
table4_new
```


`table1` is the easiest representation to work with (as seen in s12.2). (It is TIDY.) Doesn't need filtering and `rate` just adds another column to existing tibble :
```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)                  # 3.,4.
```

-----------

Q3. Recreate the plot showing change in cases over time using `table2` instead of `table1`. What do you need to do first?

A3.
[Question is referring to an example at the end of s12.2, that used `table1`]
Replacing `table1` with `table2_new` in the code. (No other changes made) :
```{r}
# Compute cases per year
table2_new %>% 
  count(year, wt = cases)
# Visualise changes over time
  ggplot(table2_new, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```
This gave the same results as with `table1` (which is same data, so ok). But it seems too easy, (and the first 2 lines aren't even needed for the plot), so probably not what the Q meant

Trying again, starting with `table2` (not `table2_new`). So can't simply use `aes(year, cases)`. Need to filter to use just the `cases` rows, and use `count` column on the y-axis :
```{r}
table2 %>% 
  filter(type == "cases") %>%
  ggplot(aes(year, count)) + 
  geom_line(aes(group = country), colour = "grey50") +          # same as before
  geom_point(aes(colour = country))                             # same as before
```


-----------

## 12. "Tidy data" [Chapter 9 hardcopy]
### 12.3.3 Exercises

Q1.
Why are `gather()` and `spread()` not perfectly symmetrical?
Carefully consider the following example:
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
# key = "year", value = "cases" ...and columns to gather are at the end (v's at start, in s12.3 teaching)
```
(Hint: look at the variable types and think about column _names_.)

Both `spread()` and `gather()` have a `convert` argument. What does it do?

A1.
The `stocks` tibble is created with 3 columns : `year`, `half`, `return`. Doing `spread()` [displays `half`, `2015`, `2016` columns] and then `gather()` [displays `half`, `year`, `return`] one may expect to return ('symmetrical') to the original tibble ...but `year` began as type `<dbl>` and is now of type `<chr>` ...(And column order differs)

'Why?' : `gather()` will convert numeric variable names to strings ...(And columns not involved in the spread/gather 'move left')

```{r}
?spread
?gather
```

Repeating, but this time with a `convert` argument :
```{r}
stocks %>% 
  spread(year, return) %>%
  gather("year", "return", `2015`:`2016`, convert = TRUE)
```
`convert=TRUE` will change a variable's class from `<chr>` to numeric/integer/logical, if it was that before the `gather()`.
(Doesn't work for factor or date.) 'Help' reads similarly for `spread()`, but I think its use is more nuanced/complicated
..."Use `convert = TRUE` to produce variables of mixed type".
In example above `year` ends up converted to `<int>`. (It started as `<dbl>`, and would have become `<chr>`)

-----------

Q2.
Why does this code fail? :
```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
```

A2.
"Error in `inds_combine(.vars, ind_list)` : Position must be between 0 and n" -- error message

* "Note that “1999” and “2000” are non-syntactic names (because they don’t start with a letter) so we have to surround them in backticks" -- s12.3.1 R4DS book
* "The tidyverse functions will interpret `1999` and `2000` without quotes as looking for the 1999th and 2000th column of the data frame" -- https://jrnold.github.io/r4ds-exercise-solutions/tidy-data.html#spreading-and-gathering

-----------

Q3.
Why does spreading this tibble fail? How could you add a new column to fix the problem? :
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

A3.
```{r}
people
spread(people, key = key, value = value)
```
"Error: Duplicate identifiers for rows (1, 3)" -- error message

* `Spread` fails because rows 1 & 3 both have “age” for “Phillip Woods”
* To fix this, could add a new column, where rows 1 & 3 have different values for a new categorical variable (eg. person who collected the data?), to distinguish them
* In this example (after fixing), the spread tibble will have `NA` for `height` in one row

-----------

Q4.
Tidy the simple tibble below. Do you need to spread or gather it? What are the variables? :
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

A4.
The variables are (`pregnant` and) `sex` and `count`. Need to `gather` it. (`key = "sex"` and `value = "count"`) :
```{r}
preg %>%
  gather(male, female, key = "sex", value = "count")
# I assume `count` is `baby_qty`, but safer to leave as `count`
```


-----------

## 12. "Tidy data" [Chapter 9 hardcopy]
### 12.4.3 Exercises

Q1.
What do the `extra` and `fill` arguments do in `separate()`?

Experiment with the various options for the following two toy datasets.
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))          # "Warning message: Too many values at 1 locations: 2". 3x3 tibble
```

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))          # "Warning message: Too few values at 1 locations: 2". 3x3 tibble
```

A1.
```{r}
?tibble
?separate
```

`extra` : controls what happens when there are too many pieces. Three valid options :

* `warn` (the default): emit a warning and drop extra values.
* `drop`: drop any extra values without a warning.
* `merge`: only splits at most 'length(into)' times

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")  # extra values not split, so “f,g” is in col "three" row 2
```


`fill` :  controls what happens when there are not enough pieces. Three valid options :

* `warn` (the default): emit a warning and fill from the right.
* `right`: fill with missing values on the right.
* `left`: fill with missing values on the left

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")    # col “one” has the <NA>, instead of col "three" (row 2)
```

-----------

Q2. Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would you set it to `FALSE`?

A2.
```{r}
?unite        # "If TRUE, remove input columns from output data frame"
?separate     # "If TRUE, remove input column from output data frame"
```
You would set it to `FALSE` if you want to **also** keep the old column(s)

-----------

Q3.
Compare and contrast `separate()` and `extract()`.
Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

A3.
```{r}
?extract
```
"Given a regular expression with capturing groups, `extract()` turns each group into a new column. If the groups don't match, or the input is `NA`, the output will be `NA`" -- Help

* `extract()` uses a regular expression to find **groups** and split into columns.
* `separate()` : relation is one:many (cols), multiple ways to split the string. Each of 3 might be most useful tool for it.
* `unite()` : relation is many:one (cols). Once cols are specified, there is only one way, (only choice is the `sep`) -- based on https://jrnold.github.io/r4ds-exercise-solutions/tidy-data.html#separating-and-uniting

NOTE : Still not sure about how `extract()` works, nor how it does/not differ from `separate()`


-----------

## 12. "Tidy data" [Chapter 9 hardcopy]
### 12.5.1 Exercises

Q1. Compare and contrast the `fill` arguments to `spread()` and `complete()`
```{r}
?spread
?complete
```

A1.

* `spread()` : `fill` sets the value. ("missing values [both explicit & implicit] will be replaced with this value")
* `complete()` : `fill` sets a value from a named list. ("for each variable supplies a single value") [implicit & explicit missing values]

-----------

Q2. What does the `direction` argument to `fill()` do? :
```{r}
?fill
```

A2.
"Direction in which to fill missing values. Currently either 'down' (the default) or 'up'" -- Help
...ie. Replace with the most recent non-missing value ("down") or the next non-missing value ("up")


-----------

## 12. "Tidy data" [Chapter 9 hardcopy]
### 12.6.1 Exercises

Q1.
In this case study I set `na.rm = TRUE` just to make it easier to check that we had the correct values. Is this reasonable?
Think about how missing values are represented in this dataset. Are there implicit missing values?
What’s the difference between an `NA` and zero?

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2.
What happens if you neglect the `mutate()` step? (`mutate(key = stringr::str_replace(key, "newrel", "new_rel"))`)

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q3.
I claimed that `iso2` and `iso3` were redundant with `country`. Confirm this claim

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q4.
For each `country`, `year`, and `sex` compute the total number of cases of TB. Make an informative visualisation of the data

A4.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

