---
title: "Chapter25"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(modelr)
# library(gapminder)                # not available in anaconda
```


## 25. "Many models" [Chapter 20 hardcopy]
### 25.2.5 Exercises

Q1.
A linear trend seems to be slightly too simple for the overall trend.

* Can you do better with a quadratic polynomial?
* How can you interpret the coefficients of the quadratic? (Hint: you might want to transform `year` so that it has mean zero.)

A1.
(I think) this refers to `lm(lifeExp ~ year, data = df)` from s25.2.2 (R4DS book) leaving some large residuals (especially for some African countries during 1980-2000 ish). A quadratic polynomial would include a 'squared' element (of `year`), and 'has mean zero' presumably suggests taking into account what `lifeExp` has been doing to-date (??)

I don't have the **gapminder** package. See [https://jrnold.github.io/r4ds-exercise-solutions/many-models.html#gapminder], which ends with :

```{r}
by_country %>% 
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance, .drop = TRUE) %>%
  ggplot(aes(continent, r.squared)) +
  geom_jitter(width = 0.5)
```


-----------

Q2.
Explore other methods for visualising the distribution of `R squared` per continent. You might want to try the **ggbeeswarm** package, which provides similar methods for avoiding overlaps as jitter, but uses deterministic methods

A2.
I don't have the **gapminder** package. See [https://jrnold.github.io/r4ds-exercise-solutions/many-models.html#gapminder] :

```{r}
by_country %>% 
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance, .drop = TRUE) %>%
  ggplot(aes(continent, r.squared)) +
  geom_beeswarm()                            # instead of `geom_jitter()`
```


-----------

Q3.
To create the last plot (showing the data for the countries with the worst model fits), we needed two steps: we created a data frame with one row per country and then semi-joined it to the original dataset.

It’s possible avoid this join if we use `unnest()` instead of `unnest(.drop = TRUE)`. How?

A3.
This refers to `bad_fit` data frame, plotted as a line chart, at the end of s25.2.4 R4DS book

An earlier step (in s25.2.4 R4DS book) towards `bad_fit` data frame tried `unnest(glance)`, but "This isn’t quite the output we want, because it still includes all the list columns. This is default behaviour when `unnest()` works on single row data frames. To suppress these columns we use `.drop = TRUE`" ...And then the book used `unnest(glance, .drop = TRUE)` instead

SO...xxxxxxxxxxx


-----------

## 25. "Many models" [Chapter 20 hardcopy]
### 25.4.5 Exercises

Q1. List all the functions that you can think of that take a atomic vector and return a list.

A1.
"Some useful functions take an atomic vector and return a list. For example, in strings [s14 R4DS book] you learned about `stringr::str_split()` which takes a character vector and returns a list of character vectors. If you use that inside `mutate`, you’ll get a list-column" -- s25.4.2 R4DS book

From [https://jrnold.github.io/r4ds-exercise-solutions/many-models.html#creating-list-columns] : "E.g. Many of the **stringr** functions"

-----------

Q2. Brainstorm useful summary functions that, like `quantile()`, return multiple values.

A2.
From [https://jrnold.github.io/r4ds-exercise-solutions/many-models.html#creating-list-columns] : "Some examples of summary functions that return multiple values are `range` and `fivenum`"

-----------

Q3.

* What’s missing in the following data frame?
* How does `quantile()` return that missing piece?
* Why isn’t that helpful here? :

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg))) %>% 
  unnest()
#> # A tibble: 15 × 2
#>     cyl     q
#>   <dbl> <dbl>
#> 1     4  21.4
#> 2     4  22.8
#> 3     4  26.0
#> 4     4  30.4
#> 5     4  33.9
#> 6     6  17.8
#> # ... with 9 more rows
```


A3.
`q` gives the `mpg` at which certain quantiles/probabilities or met for each `cyl`. The output indicates that there are 5 quantiles here, but I don't know what they are (nor how the default is calculated, if a 'numeric vector of probabilities' isn't provided)

```{r}
?quantile
```

* The data frame is missing the quantiles/probabilities. "To make useful results with unnest, you’ll also need to capture the probabilities" -- s25.4.3

* s25.4.3 has `quantile()` return 'that missing piece' by providing a 'numeric vector of probabilities' ...In the case of the data in Q3, this doesn't quite match, but replacing 0.99 with 1.00 (trial & error) would give a match :
```{r}
probs <- c(0.01, 0.25, 0.5, 0.75, 0.99)
mtcars %>% 
  group_by(cyl) %>% 
  summarise(p = list(probs), q = list(quantile(mpg, probs))) %>% 
  unnest()
```

* However, the question presumably isn't asking about trial & error (because "Why isn’t that helpful here?" wouldn't make sense)

INSTEAD, from [https://jrnold.github.io/r4ds-exercise-solutions/many-models.html#creating-list-columns], parts 2&3 are :

* `quantile()` returns these in the names of the vector :
```{r}
quantile(mtcars$mpg)
```

* Since the `unnest` function drops the names of the vector, they aren’t useful here


-----------

Q4.
What does this code do? Why might might it be useful? :
```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise_each(funs(list))
```


A4.


-----------

## 25. "Many models" [Chapter 20 hardcopy]
### 25.5.3 Exercises

Q1. Why might the `lengths()` function be useful for creating atomic vector columns from list-columns?

A1.

-----------

Q2. List the most common types of vector found in a data frame. What makes lists different?

A2.



