---
title: "Chapter28"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.2.1 Exercises

Q1.
Create one plot on the fuel economy data with customised `title`, `subtitle`, `caption`, `x`, `y`, and `colour` labels

A1.
Combining the examples from s28.2 R4DS book (which corrects/removes `paste()` code that is in the hardcopy) :

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```


-----------

Q2.
The `geom_smooth()` is somewhat misleading because the `hwy` for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model

A2.
Starting from the code in A2 and making 2 ADDitions, then revisualising :

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +        # ADD: `colour = class` for multiple lines
  geom_point(aes(colour = class)) +
  geom_smooth(method = "lm", se = FALSE) +            # ADD: `method = "lm"` for straight lines
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

Based on [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#label] (but that answer had relabelled x&y axis the wrong way around) :

```{r}
mod <- lm(hwy ~ class, data = mpg)             # model based on fuel economy for Class (Car type)
mpg %>%
  add_residuals(mod) %>%
  ggplot(aes(displ, resid)) +                  # `resid` is default column name for residuals
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Highway MPG for Cars After Subtracting Mean MPG of their Class",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy, relative to Class (mpg)"
  )
```


-----------

Q3.
Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

A3.


-----------

## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.3.1 Exercises

Q1. Use `geom_text()` with infinite positions to place text at the four corners of the plot

A1.

-----------

Q2.
Read the documentation for `annotate()`. How can you use it to add a text label to a plot without having to create a tibble?

A2.

-----------

Q3.
How do labels with `geom_text()` interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

A3.

-----------

Q4. What arguments to `geom_label()` control the appearance of the background box?

A4.

-----------

Q5.
What are the four arguments to `arrow()`? How do they work? Create a series of plots that demonstrate the most important options

A5.


-----------

## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.4.4 Exercises

Q1.
Why doesn’t the following code override the default scale?

```{r}
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_colour_gradient(low = "white", high = "red") +
  coord_fixed()
```

A1.

-----------

Q2. What is the first argument to every scale? How does it compare to `labs()`?

A2.

-----------

Q3.
Change the display of the presidential terms by:

1. Combining the two variants shown above
2. Improving the display of the y axis
3. Labelling each term with the name of the president
4. Adding informative plot labels
5. Placing breaks every 4 years (this is trickier than it seems!)

A3.

-----------

Q4.
Use `override.aes` to make the legend on the following plot easier to see

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20)
```

A4.




