---
title: "Chapter14"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(stringr)
```


## 14. "Strings" [Chapter 11 hardcopy]
### 14.2.5 Exercises

Q1.
In code that doesn’t use **stringr**, you’ll often see `paste()` and `paste0()`.
What’s the difference between the two functions? What **stringr** function are they equivalent to?
How do the functions differ in their handling of `NA`?

A1.
```{r}
?paste
```

* "`paste` converts its arguments to character strings, and concatenates them (separating them by the string given by `sep`)"
* "`paste0(..., collapse)` is equivalent to `paste(..., sep = "", collapse)`, slightly more efficiently" -- Help
* '`paste` separates strings by spaces by default, while `paste0` does not separate strings by default. As `str_c()` does not separate strings by default, it is closer in behaviour to `paste0`' -- based on https://jrnold.github.io/e4qf/strings.html

* `str_c()` : "missing values are contagious. If you want them to print as "NA", use `str_replace_na()`" -- s14.2.2
* `paste()` : "coerces `NA`, the character missing value, to "NA"" -- Help
* `str_c()` propagates `NA`. If any argument is a missing value, it returns a missing value (as do most other functions in R). However, the paste functions convert `NA` to the string "NA" and then treat it as any other character vector -- based on https://jrnold.github.io/e4qf/strings.html

-----------

Q2.
In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

A2.
```{r}
?str_c

# "Use the `sep` argument to control how they’re separated" -- s14.2.2 R4DS book
str_c("x", "y", sep = ", ")

# "To collapse a vector of strings into a single string, use `collapse`" -- s14.2.2 R4DS book
str_c(c("x", "y", "z"), collapse = ", ")
```

"`sep`	: String to insert between input vectors" -- Help
"`collapse` : Optional string used to combine input vectors into single string" -- Help

`sep` is the string inserted between arguments to `str_c()`
`collapse` is the string used to separate any elements of the character vector into a character vector of length one

-----------

Q3.
Use `str_length()` and `str_sub()` to extract the middle character from a string.
What will you do if the string has an even number of characters?

A3.
```{r}
?ceiling
```

To handle an even number of characters, choose formula to 'round' up or down. ('up' is ok with only one character)
```{r}
x <- c("a", "abc", "abcd", "abcde", "abcdef")
L <- str_length(x)
m <- ceiling(L / 2) # choosing 'floor()' here would give a zero value for the single character "a" string
str_sub(x, m, m) # [1] "a" "b" "b" "c" "c". This isn't really rounding up from middle of "abcd", as it chooses "b" etc
```
-- https://jrnold.github.io/e4qf/strings.html


-----------

Q4.
What does `str_wrap()` do? When might you want to use it?

A4.
```{r}
?str_wrap
```

"Wrap strings into nicely formatted paragraphs ... Implements the Knuth-Plass paragraph wrapping algorithm" -- Help
"wraps text so that it fits within a certain width ... For wrapping long strings of text to be typeset" -- https://jrnold.github.io/e4qf/strings.html

-----------

Q5.
What does `str_trim()` do? What’s the opposite of `str_trim()`?

A5.
```{r}
?str_trim
```

"Trim whitespace from start and end of string" -- Help
`str_pad()` adds whitespace to each side

-----------

Q6.
Write a function that turns (e.g.) a vector `c("a", "b", "c")` into the string `a, b, and c`.
Think carefully about what it should do if given a vector of length 0, 1, or 2

A6.
We haven't covered functions yet? [Chapter 19]. But this is for 2nd part :
length 0 : ""
length 1 : eg. "a"
length 2 : eg. "a and b"
length 3 : "a, b, and c"


```{r}
# However, from https://jrnold.github.io/e4qf/strings.html#exercises-23 :
str_commasep <- function(x, sep = ", ", last = ", and ") {
  if (length(x) > 1) {
    str_c(str_c(x[-length(x)], collapse = sep),
          x[length(x)],
          sep = last)
  } else {
    x
  }
}
# I don't really understand how this is working, particularly code inside the first {}
```

XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.3.1.1 Exercises

Q1.
Explain why each of these strings <in a regexp> don’t match a `\` <within strings you're searching> : `"\"`, `"\\"`, `"\\\"`

A1.
If \ is used as an escape character in regular expressions, how do you *match* a literal \? Well you need to escape it, creating the regular expression \\. To create that regular expression, you need to use a string, which also needs to escape \. That means to match a literal \ you need to write "\\\\" — you need four backslashes to match one! -- s14.3.1

* "\" : will escape the next character in the R string
* "\\" : will resolve to \ in the regexp, which will escape the next character in the regexp
* "\\\" : The first two will resolve to a literal backslash in the regexp, the third will escape the next character. So in the regexp, this will escape some escaped character. <??? EH?>
-- https://jrnold.github.io/r4ds-exercise-solutions/strings.html#matching-patterns-and-regular-expressions

XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2. How would you match <to> the sequence `"'\` <if it exists within a series of strings you're searching> ?

A2.
"To include a literal single or double quote in a string you can use \ to “escape” it" -- s14.2 R4DS book

"Beware that the printed representation of a string is not the same as string itself, because the printed representation shows the escapes. To see the raw contents of the string, use `writeLines()`" -- s14.2 R4DS book

```{r}
?'"' # in order to see list of Character Constants (like \n, \t, etc) on the Help page
?str_view # "`str_view` shows the first match; `str_view_all` shows all the matches"

x <- '\"\'\\'
x                  # shows [1] "\"'\\"
writeLines(x)      # shows "'\
# Does that answer the question?
```

-----------

Q3.
What patterns will the regular expression `\..\..\..` match? How would you represent it as a string?

A3.
`......` : This is a literal dot followed by a 'wildcard', 3 times
It will match any patterns that are a dot followed by any character, 3 times

```{r}
y <- '\\..\\..\\..'
y                  # shows [1] "\\..\\..\\.."
writeLines(y)      # shows \..\..\..
```


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.3.2.1 Exercises

Q1.
How would you <write a regexp to> match the literal string `"$^$"` <if it exists within strings you're searching> ?

A1.
"To force a regular expression to only match a complete string, anchor it with both `^` and `$`" -- s14.3.2

```{r}
z <- "^\\$\\^\\$$" # Able to take a guess at this, but I still need to check/adjust using `writeLines()` and `str_view()`
z                  # shows [1] "^\\$\\^\\$$"
writeLines(z)      # shows ^\$\^\$$

str_view(c("$^$", "pp$^$pp", "$^$qq", "rr$^$"), "^\\$\\^\\$$")  # as a test
```

-----------

Q2. Given the corpus of common words in `stringr::words`, create regular expressions that find all words that:
Q2.1 Start with “y”.
Q2.2 End with “x”
Q2.3 Are exactly three letters long. (Don’t cheat by using `str_length()`!)
Q2.4 Have seven letters or more.
Since this list is long, you might want to use the `match` argument to `str_view()` to show only the matching or non-matching words

```{r}
?str_view
stringr::words

str_view(stringr::words, "^y", match = TRUE)         # A2.1 (6 words)
str_view(stringr::words, "x$", match = TRUE)         # A2.2 (4 words)
str_view(stringr::words, "^...$", match = TRUE)      # A2.3 ? (16 words start a/b)
str_view(stringr::words, "^.......", match = TRUE)   # A2.4 ? (>=16 words start a)
```


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.3.3.1 Exercises

Q1. Create regular expressions to find all words that:
Q1.1 Start with a vowel
Q1.2 That only contain consonants. (Hint: thinking about matching “not”-vowels)
Q1.3 End with ed, but not with eed
Q1.4 End with ing or ise

A1.1
```{r}
str_view(stringr::words, "^(a|e|i|o|u)", match = TRUE)

str_view(stringr::words, "^[aeiou]", match = TRUE)     # Better/clearer than above, to choose from single letters
```

A1.2
```{r}
str_view(stringr::words, "^[^aeiou]", match = TRUE)    # Starts with a consonant. HOW DO I DO THAT FOR THE REST?
```
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

A1.3
```{r}
str_view(stringr::words, "[^e]ed$", match = TRUE)     # No 2-letter word "ed" is in stringr::words, so this is ok
```


A1.4
```{r}
str_view(stringr::words, "ing$|ise$", match = TRUE)   # Or could use "i(ng|se)$", which is shorter (but less clear?)
```

-----------

Q2.
Empirically verify the rule “i before e except after c”

A2.
```{r}
str_view(stringr::words, "([^c]ie|cei)", match = TRUE)     # 14 words match the rule
str_view(stringr::words, "([^c]ei|cie)", match = TRUE)     # But 3 (ie. >0) words here show it to actually be wrong
```

-----------

Q3.
Is “q” always followed by a “u”?

A3.
```{r}
str_view(stringr::words, "q[^u]", match = TRUE)            # Zero words match, so 'yes' (from this list of words)
```

-----------

Q4.
Write a regular expression that matches a word if it’s probably written in British English, not American English.

A4.
```{r}
str_view(stringr::words, "(our$|ise$|^bloke$|^brilliant$|ntre|^chap$|^quid$)", match = TRUE)
```
Some false +ves, and came from looking through the list of `stringr::words` (though I did think of 'our' & 'ise')


-----------

Q5.
Create a regular expression that will match telephone numbers as commonly written in your country

A5.
```{r}
x <- c("0117 912 3456", "01179123456")
str_view(x, "\\d\\d\\d\\d \\d\\d\\d \\d\\d\\d\\d")       # matches the first one, not the second
str_view(x, "\\d\\d\\d\\d\\s\\d\\d\\d\\s\\d\\d\\d\\d")   # alternate ('better') way, using '\s' for whitespace
```


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.3.4.1 Exercises
  
Q1.
Describe the equivalents of `?`, `+`, `*` in `{m,n}` form

A1.
`?` : 0 or 1    -- s14.3.4 : `{,1}`, matching at most 1
`+` : 1 or more -- s14.3.4 : `{1,}`, matching 1 or more
`*` : 0 or more -- s14.3.4 : "There is no direct equivalent of * in {m,n} form since there are no bounds on the matches:
#                              it can be 0 up to infinity matches" -- https://jrnold.github.io/e4qf/strings.html
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX CAN WE USE {0,} or {,} ?

-----------

Q2.
Describe in words what these regular expressions match:
(read carefully to see if I’m using a regular expression or a string that defines a regular expression)
Q2.1 `^.*$`
Q2.2 `"\\{.+\\}"`
Q2.3 `\d{4}-\d{2}-\d{2}`
Q2.4 `"\\\\{4}"`

A2.1 `.*`             : Would match any string (of any length)
A2.2 `\{.+\}`         : Would match curly braces around >= 1 character
A2.3 eg. 1111-22-33   : 4 digits dash 2 digits dash 2 digits. Could be a date, with Year first
A2.4 `\\{4}`          : 4 backslashes

-----------

Q3.
Create regular expressions to find all words that:
Q3.1 Start with three consonants
Q3.2 Have three or more vowels in a row
Q3.3 Have two or more vowel-consonant pairs in a row

```{r}
str_view(stringr::words, "^[^aeiou]{3}", match = TRUE)              # A3.1
str_view(stringr::words, "[aeiou]{3,}", match = TRUE)               # A3.2
str_view(stringr::words, "([aeiou][^aeiou]){2,}", match = TRUE)     # A3.3
```

-----------

Q4. Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner
A4. (Did the 1st one ok. Then didn't know what \1 was, in the 2nd and 3rd. And gave up)


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.3.5.1 Exercises

Q1.
Describe, in words, what these expressions will match:
Q1.1 `(.)\1\1`
Q1.2 `"(.)(.)\\2\\1"`
Q1.3 `(..)\1`
Q1.4 `"(.).\\1.\\1"`
Q1.5 `"(.)(.)(.).*\\3\\2\\1"`

A1.1 Any character (the same one) 3 times in a row. (eg. xxx)
A1.2 character1, character2, character2, character1. (eg. xyyx)
A1.3 Two characters, repeated again. (eg. abab)
A1.4 character1, character_any, character1, character_any, character1. (eg. abaca)
A1.5 Three characters, then some characters, then the original three in reverse. (eg. xyz45826zyx)

-----------

Q2.
Construct regular expressions to match words that:
Q2.1 Start and end with the same character
Q2.2 Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
Q2.3 Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

```{r}
str_view(stringr::words, "^(.).*\\1$", match = TRUE)           # A2.1
str_view(stringr::words, "(..).*\\1", match = TRUE)            # A2.2
str_view(stringr::words, "(.).*\\1.*\\1", match = TRUE)        # A2.3
```


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.4.2 Exercises

Q1.
For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple `str_detect()` calls:
Q1.1 Find all words that start or end with x
Q1.2 Find all words that start with a vowel and end with a consonant
Q1.3 Are there any words that contain at least one of each different vowel?

# A1.1
str_view(stringr::words, "^x|x$", match = TRUE)                     # (None starts with x.) End: box, sex, six, tax
stringr::words[str_detect(stringr::words, "^x|x$")]                 # Alternative. Same results in Console (not Viewer)
str_subset(stringr::words, "^x|x$")                                 # Alternative. Same results in Console (not Viewer)

start_x <- str_detect(stringr::words, "^x")
end_x <- str_detect(stringr::words, "x$")
stringr::words[start_x | end_x]                                     # Combination. Same results in Console


# A1.2
str_view(stringr::words, "^[aeiou].*[^aeiou]$", match = TRUE)       # In Viewer
stringr::words[str_detect(stringr::words, "^[aeiou].*[^aeiou]$")]   # Alternative. In Console
str_subset(stringr::words, "^[aeiou].*[^aeiou]$")                   # Alternative. In Console

start_vowel <- str_detect(stringr::words, "^[aeiou]")
end_consonant <- str_detect(stringr::words, "[^aeiou]$")
stringr::words[start_vowel & end_consonant]                         # Combination. Same results in Console


# A1.3
str_view(stringr::words, "([aeiou]).*\\1.*\\1.*\\1.*\\1", match = TRUE)
# None has 5 vowels, so none has 5 different. (WRONG - This code looks for the *same* vowel 5 times)
str_view(stringr::words, "[aeiou].*[aeiou].*[aeiou].*[aeiou].*[aeiou]", match = TRUE)
# Returns 8 words with 5 vowels. By *inspection*, none has 5 different. ('associate', 'colleague', 'appropriate' have 4)

contains_a <- str_detect(words, "a")
contains_e <- str_detect(words, "e")
contains_i <- str_detect(words, "i")
contains_o <- str_detect(words, "o")
contains_u <- str_detect(words, "u")
words[contains_a & contains_e & contains_i & contains_o & contains_u]
# Seemed to work ok (no errors). No matches (as expected, but doesn't prove that code works)
# Replaced 'words' with "baeiout", and code worked (matched) -- Suggested by https://jrnold.github.io/e4qf/strings.html

-----------

Q2.
What word has the highest number of vowels? What word has the highest proportion of vowels?
(Hint: what is the denominator?)

A2.
```{r}
words[str_detect(words, "[aeiou].*[aeiou].*[aeiou].*[aeiou].*[aeiou]")]              # 5 vowels. re. A1.3
words[str_detect(words, "[aeiou].*[aeiou].*[aeiou].*[aeiou].*[aeiou].*[aeiou]")]     # BUT 6 vowels. No match

proportion_vowels <- str_count(words, "[aeiou]") / str_length(words)        # proportion, for each word
words[which(proportion_vowels == max(proportion_vowels))]                   # max is 100%, when the word is a vowel
# -- https://jrnold.github.io/e4qf/strings.html
# `str_length()` was in s14.2.1 so that's ok ...BUT we haven't coverered `which()` or `max()` in the book

proportion_vowels <- str_count(words, "[aeiou]") / str_length(words)        # proportion, for each word
words[proportion_vowels == 1]                                               # max is 100%, when the word is a vowel
```


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.4.3.1 Exercises

Q1.
In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour.
Modify the regex to fix the problem

A1.
Question is referring to s14.4.3
regex could have whitespace [\s] each side of each colour word. But I got confused which bits of s14.4.3 code did what.
https://jrnold.github.io/e4qf/strings.html : "Add the \b before and after the pattern", which is similar but different to \s. ("You can also match the boundary between words with \b" -- s14.3.2). Copied from there (but edited `more2`) :
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")       # This is the row that answers the question
more2 <- sentences[str_count(sentences, colour_match2) > 1]
str_view_all(more2, colour_match2, match = TRUE)
```

-----------

Q2.
From the Harvard sentences data, extract:
Q2.1 The first word from each sentence
Q2.2 All words ending in ing
Q2.3 All plurals

```{r}
?str_extract
str_extract(stringr::sentences, "^\\b(.+)\\b")      # A2.1 Why does this return whole sentence? - doesn't \b find a word?
str_extract(stringr::sentences, "^\\b.+\\b")        # Returned whole sentence

str_extract(stringr::sentences, "^\\b[a-z]+\\b")
str_extract(stringr::sentences, "^\\b[a-z]+?\\b")

str_extract(stringr::sentences, "^\\b(.+)?\\b")     # Returned whole sentence ***
str_extract(stringr::sentences, "^\\b(.+)\\b?")     # Error
str_extract(stringr::sentences, "^\\b.+?\\b")       # THIS ONE WORKED         ***
str_extract(stringr::sentences, "^\\b.+\\b?")       # Error
str_extract(stringr::sentences, "^\\b.?\\b")
str_extract(stringr::sentences, "^\\b.+?")          # Returned 1st letter
str_extract(stringr::sentences, "^.+?\\b")          # THIS ONE WORKED
str_extract(stringr::sentences, "^.+?")             # Returned 1st letter

str_extract(stringr::sentences, "[A-Za-z]+\\w")     # THIS ONE WORKED
str_extract(stringr::sentences, "[A-Za-z]+")        # THIS ONE WORKED         ***

str_extract(stringr::sentences, ".+?")              # Returned 1st letter
str_extract(stringr::sentences, "(.+)?")            # Returned whole sentence
str_extract(stringr::sentences, "^\\b[A-Za-z]+\\b") # THIS ONE WORKED
```


A2.2
A2.3
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
Answers to Q2 at https://jrnold.github.io/e4qf/strings.html use things we haven't covered in the book ...`[a-zA-Z]`, `unique()`, `unlist()`


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.4.4.1 Exercises

Q1.
Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2.
Find all contractions. Separate out the pieces before and after the apostrophe

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.4.5.1 Exercises

Q1.
Replace all forward slashes in a string with backslashes

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2.
Implement a simple version of `str_to_lower()` using `replace_all()`

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q3.
Switch the first and last letters in words. Which of those strings are still words?

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.4.6.1 Exercises

Q1.
Split up a string like "apples, pears, and bananas" into individual components

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2. Why is it better to split up by `boundary("word")` than `" "`?

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q3.
What does splitting with an empty string ("") do? Experiment, and then read the documentation

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.5.1 Exercises

Q1.
How would you find all strings containing `\` with `regex()` vs. with `fixed()`?

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


-----------

Q2.
What are the five most common words in sentences?

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

## 14. "Strings" [Chapter 11 hardcopy]
### 14.7.1 Exercises

Q1.
Find the **stringi** functions that:
Q1.1 Count the number of words
Q1.2 Find duplicated strings
Q1.3 Generate random text

A1.
A1.1 `stri_count_words`
A1.2 `stri_duplicated`
A1.3 `stri_rand_strings` generates random strings. (`stri_rand_lipsum` generates lorem ipsum text)

-----------

Q2.
How do you control the language that `stri_sort()` uses for sorting?

A2.
```{r}
?stri_sort
```
Use the `locale` argument to the `opts_collator` argument. [See also `stri_opts_collator()`]


