---
title: "Chapter03"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.2.4 Exercises

Q1. Run `ggplot(data = mpg)`. What do you see?

A1.
```{r}
ggplot(data = mpg)
```
Grey background. No axes or data visible

-----------

Q2. How many rows are in `mpg`? How many columns?

A2.
```{r}
mpg
```
234 rows x 11 columns

-----------

Q3. What does the `drv` variable describe? Read the help for `?mpg` to find out

A3.
```{r}
?mpg
```
`f` = front-wheel drive, `r` = rear wheel drive, `4` = 4wd

-----------

Q4. Make a scatterplot of `hwy` vs `cyl` ...[ie. y v's x]

A4.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

-----------

Q5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

A5.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))
```
The plot has 2 categorical variables, (but shows no dependency), with an unknown number of points at many intersections (ie. overlap)


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.3.1 Exercises

Q1. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

A1. Needs to be outside the `aes()` to work like that. (Inside, it just creates a 'True' condition, but with no say over what colour represents it in the plot)

-----------

Q2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

A2.
```{r}
?mpg
mpg
```
* Categorical : `<chr>` : `manufacturer`; `model`; `trans`; `drv`; `fl`; `class` ...6
* Continuous : `<dbl>` or `<int>` : `displ`; `year`; `cyl`; `cty`; `hwy` ...5

-----------

Q3.
Map a continuous variable to `color`, `size`, and `shape` ...[I initially misunderstood the requirement, and mapped them all in one plot]
How do these aesthetics behave differently for categorical vs. continuous variables? ...[eg. `drv` `<chr>`; `cty` `<int>`]

A3.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv))
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = drv))
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = drv))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = cty))
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```
`ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = cty))` # "Error: A continuous variable cannot be mapped to shape""

ANSWER [categorical v's continuous] :

* `color` : different hues v's tints & shades of a single hue
* `size` : allocation of size to variable is arbitrary (warning) v's relates to size of variable
* `shape` : allocation of shape to variable is arbitrary v's error (as 'meaningless')

-----------

Q4. What happens if you map the same variable to multiple aesthetics?

A4.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ, size = hwy))
```
The plot does use all the aesthetics. But there is redundancy, which may be confusing (as a reader expects that an aesthetic is added for extra insight)

-----------

Q5. What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

A5.
```{r}
?geom_point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)
```
"For shapes that have a border (like 21), you can colour the inside and outside separately. Use the `stroke` aesthetic to modify the width of the border"

-----------

Q6. What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)` ?

A6.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```
Expression is True/False for each point, splitting `displ` into 2 sets with different colours ...(Remember Q1 with `color = "blue"`)


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.5.1 Exercises

Q1. What happens if you facet on a continuous variable?

A1.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_wrap(~ displ)
```
A cell is created for every value of the continuous variable. (Unlikely to be the most insightful choice)

-----------

Q2. What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```

A2.

* 1st plot maps combinations of 2 categorical variables. No point at an intersection means the data doesn't contain that combination
* 2nd plot has an empty cell when 'the data doesn't contain that combination'

-----------

Q3. What plots does the following code make? What does `.` do?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

A3.

* 1st plot has long horizontal cells (3 rows, no cols)
* 2nd plot has tall vertical cells (0 rows, 4 cols)
* `.` acts as a placeholder in the script (when facet_grid is missing a variable), and causes no faceting in that dimension

-----------

Q4.
Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages?
How might the balance change if you had a larger dataset?

A4.

* Adv : reduces qty of points in a cell, so less overlapping and easier to spot different trends
* Disadv : harder to judge location of points in a cell relative to other cells (ie. overall picture). Squeezes the size of a cell
* Larger dataset has more overlapping, unless faceting used. (Though if the categorical variables used for faceting have many values, the cells will be tiny)

-----------

Q5. 
Read `?facet_wrap`. What does `nrow` do? What does `ncol` do?
What other options control the layout of the individual panels?
Why doesn’t `facet_grid()` have `nrow` and `ncol` argument?

A5.
```{r}
?facet_wrap
```
"`facet_wrap` wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than `facet_grid` because most displays are roughly rectangular" -- Help

* `nrow` breaks a single row of cells into the specified number of rows. Similarly `ncol` specifies number of cols to have in the plot
* `as.table` [= TRUE or FALSE] determines the starting facet to begin filling the plot. And `dir` [= "h" or "v"] determines the starting direction for filling in the plot
* `facet_grid()` has its x and y dimensions determined by its arguments (number of unique values of the 2 variables)

-----------

Q6. When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

A6. PC displays are usually landscape format, so a reader can distinguish many columns more easily than many rows (and dependent variable is usually y-axis)


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.6.1 Exercises

Q1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

A1. `geom_line()`, `geom_boxplot()`, `geom_histogram()`, `geom_area()`

-----------

Q2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

A2. Ok

-----------

Q3.
What does `show.legend = FALSE` do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

A3.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(show.legend = FALSE) +
  geom_smooth(show.legend = FALSE, se = FALSE)
```

* Plot does not have a legend. Removing 'FALSE' will add a legend (because default is 'TRUE')
* 'FALSE' used earlier, to 'expand' the size/proportion of the plot used by data

-----------

Q4.What does the `se` argument to `geom_smooth()` do?

A4.
```{r}
?geom_smooth
```
"display confidence interval around smooth? (TRUE by default, see level to control" ...ie. It adds error bands to a line

-----------

Q5. Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))
```

A5.
They will look the same, because the data & mapping settings are the same in each.
(The settings cascade down ...ie. are inherited, but are superceded by local settings)

-----------

Q6. Recreate the R code necessary to generate the following graphs.

A6.1
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE)
```


A6.2
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(mapping = aes(group = drv), color = "blue", se = FALSE)
```


A6.3
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


A6.4
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```


A6.5
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```


A6.6
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(color = "white", stroke = 2) +
  geom_point(mapping = aes(color = drv))
```
* Note the white points are in the earlier layer, with the coloured points then on top
* Instead of `stroke=2` could have used `size=4` (from seeing someone else's answer)
* Actually, why did `stroke` work? - does it not need to be with a shapetype that can have color+fill? (re. 3.3.1 Exercises Q5 and its border)


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.7.1 Exercises

Q1.
What is the default geom associated with `stat_summary()`?
How could you rewrite the previous plot to use that geom function instead of the stat function?

A1.
```{r}
?stat_summary
?geom_pointrange

# 'previous plot' in the book text, before 3.7.1 Exercises :
ggplot(data = diamonds) +
  stat_summary(mapping = aes(x = cut, y = depth), fun.ymin = min, fun.ymax = max, fun.y = median)
```

Default geom for `stat_summary` is `geom_pointrange`. But default stat for `geom_pointrange` is `identity` (which is NOT what we want), so explicitly specify `geom_pointrange(stat = "summary")` [see below] :
```{r}
ggplot(data = diamonds) +
  geom_pointrange(mapping = aes(x = cut, y = depth), stat = "summary", fun.ymin = min, fun.ymax = max, fun.y = median)
```

-----------

Q2. What does `geom_col()` do? How is it different to `geom_bar()`?

A2.
```{r}
?geom_col
?geom_bar
```

* "`geom_bar` uses `stat_count` by default: it counts the number of cases at each x position.
* `geom_col` uses `stat_identity`: it leaves the data as is" -- Help
* [However, `geom_bar(stat = "identity")` and `geom_col()` are equivalent]

-----------

Q3.
Most geoms and stats come in pairs that are almost always used in concert.
Read through the documentation and make a list of all the pairs. What do they have in common?

A3.
http://ggplot2.tidyverse.org/reference/ [lists lots of pairs, after this quote] :
"A layer combines data, aesthetic mapping, a geom (geometric object), a stat (statistical transformation), and a position adjustment. Typically, you will create layers using a geom_ function, overriding the default position and stat if needed"

?

-----------

Q4. What variables does `stat_smooth()` compute? What parameters control its behaviour?

A4.
```{r}
?stat_smooth
```
* Four : `y` [predicted value]; `ymin` [lower pointwise confidence interval around the mean]; `ymax` [upper pointwise confidence interval around the mean]; `se` [standard error]
* Most importantly : `method` [smoothing method (function) to use]; `se` [whether to display confidence interval (TRUE default)]; `level` [level of confidence interval to use (0.95 default)]

-----------

Q5.
In our proportion bar chart, we need to set `group = 1`. Why?
In other words what is the problem with these two graphs?
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop..)) # 'prop' is 1.00 for each value of 'cut'

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..)) # 'prop' is 7 for each value of 'cut' (7 colors)
```

A5.
We've relied on other variables to change aggregation (eg. `color = drv` instead of a separate `group = drv`)
cf. s3.6 : "It is convenient to rely on this feature because the group aesthetic by itself does not add a legend or distinguishing features to the geoms".

But `..prop..` means proportion of group, and that will be 100% unless we force it otherwise.
`group=1` means everything is in the same group (named '1'), rather than in different groups (named Fair, Good, etc ...if group=cut) ...Imagine a field called 'group', where every row has the value '1' there.
group=2 or group=X would have the same effect. [?? Don't know why it has to be inside aes(), since it's a constant ??]
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = 1)) # Note : chart isn't coloured, and displays same as the one above

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = 1)) +
  facet_wrap(~ color)
```


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.8.1 Exercises
  
Q1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()
```

A1.

* Many of the data points overlap (overplotting).
* We can 'jitter' the position of points, to improve the overall visualization (in this case)
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_jitter()
```

-----------

Q2. What parameters to `geom_jitter()` control the amount of jittering?

A2.
```{r}
?geom_jitter
```
`width` and `height`...

"Amount of vertical and horizontal jitter. The jitter is added in both positive and negative directions, so the total spread is twice the value specified here.

If omitted, defaults to 40% of the resolution of the data: this means the jitter values will occupy 80% of the implied bins. Categorical data is aligned on the integers, so a width or height of 0.5 will spread the data so it's not possible to see the distinction between the categories" -- Help

-----------

Q3. Compare and contrast `geom_jitter()` with `geom_count()`

A3.
```{r}
?geom_count

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_count()
```
"This is a variant `geom_point` that counts the number of observations at each location, then maps the count to point area".

`geom_count()` makes larger points where there is overplotting, so the number of visible points is equal to `geom_point()`. (Whether to use this or `geom_jitter` will vary with circumstances and personal preference)

-----------

Q4.
What’s the default position adjustment for `geom_boxplot()`?
Create a visualisation of the `mpg` dataset that demonstrates it

A4.
```{r}
?geom_boxplot
```
`position = "dodge"` is the default. (cf. fill, identity, jitter, stack)

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, color = drv)) +
  geom_boxplot(position = "dodge")
```


-----------

## 3. "Data visualisation" [Chapter 1 hardcopy]
### 3.9.1 Exercises

Q1. Turn a stacked bar chart into a pie chart using `coord_polar()`

A1.
```{r}
?coord_polar # A pie chart = stacked bar chart + polar coordinates
# pie <- ggplot(mtcars, aes(x = factor(1), fill = factor(cyl))) +
#  geom_bar(width = 1)
# pie + coord_polar(theta = "y")
```

```{r}
ggplot(data = mpg, mapping = (aes(x = 1, fill = class))) + # Need x. x=1 gave a single stack, centred on '1'. x=factor(1) would get rid of extra scaling marks
  geom_bar(position = "stack")
```

```{r}
ggplot(data = mpg, mapping = (aes(x = 1, fill = class))) + # x=factor(1) would get rid of extra scaling marks
  geom_bar(position = "stack") +
  coord_polar(theta = "y") # theta="y" gave pie segments, (instead of rings, if theta="x" or left blank)
```

-----------

Q2. What does `labs()` do? Read the documentation

A2.
```{r}
?labs
```
`labs()` is a shortcut function to add labels to different scales ...title, subtitle, axes, and caption

-----------

Q3. What’s the difference between `coord_quickmap()` and `coord_map()`?

A3.
```{r}
?coord_map
```
* "`coord_quickmap()` sets the aspect ratio correctly for maps" -- R4DS text
* "`coord_map` projects a portion of the earth, which is approximately spherical, onto a flat 2D plane
using any projection defined by the `mapproj` package. Map projections do not, in general, preserve
straight lines, so this requires considerable computation. `coord_quickmap` is a quick approximation
that does preserve straight lines. It works best for smaller areas closer to the equator" -- Help

-----------

Q4. What does the plot below tell you about the relationship between city and highway mpg?
Why is `coord_fixed()` important? What does `geom_abline()` do?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed()
```

A4.
Relationship is positive and approximately linear. Slightly better highway mpg than city mpg

* `coord_fixed()` draws equal intervals on \(x\) & \(y\) axes, so more easily comparable (`abline` is at 45 degrees)
* `geom_abline()` draws a line that, by default, has an intercept of 0 and slope of 1


-----------

Comment & code from @aaron_g in (R4DS Slack) board=02_week on 15/9/17 06:25, in reply to query from @darh78.
Not one of the exercises, but saved here in case useful in future. (Slope charts in a lattice) :

"i think this kind of graph is much more informative (maybe this is what you had in mind to begin with)"
```{r}
mpg1 <- mpg %>% 
  mutate(unique = rownames(mpg)) %>%
  gather("road_type" ,"MPG", c("cty", "hwy"))

ggplot(data = mpg1,mapping = aes(x = road_type, y = MPG, color=drv ,group=unique)) +
  geom_point() +
  facet_wrap(~manufacturer, nrow=3)+
  geom_line(alpha=0.5) +
  theme_linedraw(base_size = 14)
```

