---
title: "Chapter19"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 19. "Functions" [Chapter 15 hardcopy]
### 19.2.1 Practice

Q1.

* Why is `TRUE` not a parameter to `rescale01()`?
* What would happen if `x` contained a single missing value, and `na.rm` was `FALSE`?

A1.
This refers to a comment in s19.2 R4DS book, (where the `rescale01` function later gets defined), after code that includes `na.rm = TRUE` : "This code only has one input: `df$a`. (If you’re surprised that `TRUE` is not an input, you can explore why in the exercise below.)"

* `TRUE` not a parameter to `rescale01()` because XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX

* if `x` contained a single missing value, and `na.rm` was `FALSE` :
```{r}
# "it’s a good idea to check your function with a few different inputs" -- s19.2 R4DS book :
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)             # na.rm = TRUE
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(1, 2, 3, NA, 5))                # [1] 0.00 0.25 0.50   NA 1.00

# function amended, then retested with the same input as above :
rescale01 <- function(x) {
  rng <- range(x, na.rm = FALSE)            # `na.rm = FALSE`
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(1, 2, 3, NA, 5))                # [1] NA NA NA NA NA
```

"If there were any NA values, and na.rm = FALSE, then the function would return NA" -- [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function]


-----------

Q2.
In the second variant of `rescale01()`, infinite values are left unchanged.
Rewrite `rescale01()` so that `-Inf` is mapped to 0, and `Inf` is mapped to 1

A2.

-----------

Q3.
Practice turning the following code snippets into functions. Think about what each function does.
What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?
```{r}
mean(is.na(x))                                       # Q3.1
x / sum(x, na.rm = TRUE)                             # Q3.2
sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)          # Q3.3
```

A3.
```{r}
A3.1
A3.2
A3.3
```


-----------

Q4.
Follow [http://nicercode.github.io/intro/writing-functions.html] to write your own functions to compute the variance and skew of a numeric vector

A4.

-----------

Q5.
Write `both_na()`, a function that takes two vectors of the same length and returns the number of positions that have an `NA` in both vectors

A5.

-----------

Q6.
What do the following functions do? Why are they useful even though they are so short?
```{r}
is_directory <- function(x) file.info(x)$isdir        # Q6.1
is_readable <- function(x) file.access(x, 4) == 0     # Q6.2
```


A6.1
A6.2

-----------

Q7.
Read the complete lyrics [https://en.wikipedia.org/wiki/Little_Bunny_Foo_Foo] to “Little Bunny Foo Foo”. There’s a lot of duplication in this song. Extend the initial piping example to recreate the complete song, and use functions to reduce the duplication

A7.


