---
title: "Chapter24"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)

library(nycflights13)
library(lubridate)
```


## 24. "Model building" [Chapter 19 hardcopy]
### 24.2.3 Exercises

Q1. In the plot of `lcarat` vs. `lprice`, there are some bright vertical strips. What do they represent?

A1.

This refers to a plot in s24.2.1 of R4DS book, after we "Log-transform the `carat` and `price` variables". And 'bright' is a function of 'count' in `geom_hex()`

"`log2` computes binary (i.e., base 2) logarithms" -- Help. "The log-transformation is particularly useful here because it makes the pattern linear, and linear patterns are the easiest to work with" -- R4DS book

The strips represent what we've noted before : an uneven distribution of diamond size, with the strips at 'human friendly' numbers (round numbers or fractions). (`lcarat`=0 means `carat`=1)


-----------

Q2. If `log(price) = a_0 + a_1 * log(carat)`, what does that say about the relationship between `price` and `carat`?

A2.

"`log` computes logarithms, by default natural logarithms" -- Help. [ie. `log(x)` is equivalent to `log(x, base = exp(1))`]

"An 1% increase in `carat` is associated with an `a_1`% increase in `price`" -- [https://jrnold.github.io/r4ds-exercise-solutions/model-building.html#why-are-low-quality-diamonds-more-expensive]

-----------

Q3. Extract the diamonds that have very high and very low residuals. Is there anything unusual about these diamonds? Are they particularly bad or good, or do you think these are pricing errors?

A3.

From s24.2.1 R4DS book : "To interpret the y axis [`lresid`], we need to think about what the residuals are telling us, and what scale they are on. A residual of -1 indicates that `lprice` was 1 unit lower than a prediction based solely on its weight. 2^-1 is 1/2, points with a value of -1 are half the expected price, and residuals with value 1 are twice the predicted price"

From s24.2.2 R4DS book (with 4 predictors now) : "This plot indicates that there are some diamonds with quite large residuals - remember a residual of 2 indicates that the diamond is 4x the price that we expected. It’s often useful to look at unusual values individually" ... [Some code to `filter()` residuals `> 1`] ... "Nothing really jumps out at me here, but it’s probably worth spending time considering if this indicates a problem with our model, or if there are errors in the data. If there are mistakes in the data, this could be an opportunity to buy diamonds that have been priced low incorrectly"

From [https://jrnold.github.io/r4ds-exercise-solutions/model-building.html#why-are-low-quality-diamonds-more-expensive] : "This was already discussed in the text. I don’t see anything either"


-----------

Q4.
Does the final model, `mod_diamonds2`, do a good job of predicting diamond prices?
Would you trust it to tell you how much to spend if you were buying a diamond?

A4.
I don't really follow this, but see [https://jrnold.github.io/r4ds-exercise-solutions/model-building.html#why-are-low-quality-diamonds-more-expensive] for brief calculations, followed by a conclusion of ['cleaned up']...


'The average squared and absolute errors are 2^0.192 (which is 1.1423) and 2^0.149 (which is 1.1088), so on average, the error is ±10-15%. And the 95% range of residuals [by averaging results of 0.975 and 0.025 ??] is about 2^0.37 (which is 1.2924), so within ±30%. This doesn’t seem terrible to me'


-----------

## 24. "Model building" [Chapter 19 hardcopy]
### 24.3.5 Exercises

Q1.

* Use your Google sleuthing skills to brainstorm why there were fewer than expected flights on Jan 20, May 26, and Sep 1. (Hint: they all have the same explanation.)
* How would these days generalise to another year?

A1.


-----------

Q2.
What do the three days with high positive residuals represent? How would these days generalise to another year?

daily %>% 
  top_n(3, resid)
#> # A tibble: 3 × 5
#>         date     n  wday resid   term
#>       <date> <int> <ord> <dbl> <fctr>
#> 1 2013-11-30   857   Sat 112.4   fall
#> 2 2013-12-01   987   Sun  95.5   fall
#> 3 2013-12-28   814   Sat  69.4   fall

A2.


-----------

Q3.

* Create a new variable that splits the `wday` variable into terms, but only for Saturdays,
i.e. it should have `Thurs`, `Fri`, but `Sat-summer`, `Sat-spring`, `Sat-fall`.
* How does this model compare with the model with every combination of `wday` and `term`?

A3.
```{r}

```


-----------

Q4.
Create a new `wday` variable that combines the day of week, term (for Saturdays), and public holidays. What do the residuals of that model look like?

A4.
```{r}

```


-----------

Q5.

* What happens if you fit a day of week effect that varies by month (i.e. `n ~ wday * month`)?
* Why is this not very helpful?

A5.


-----------

Q6.

* What would you expect the model `n ~ wday + ns(date, 5)` to look like?
* Knowing what you know about the data, why would you expect it to be not particularly effective?

A6.

-----------

Q7.
We hypothesised that people leaving on Sundays are more likely to be business travellers who need to be somewhere on Monday. Explore that hypothesis by seeing how it breaks down based on distance and time: if it’s true, you’d expect to see more Sunday evening flights to places that are far away

A7.
```{r}

```


-----------

Q8.
It’s a little frustrating that Sunday and Saturday are on separate ends of the plot.
Write a small function to set the levels of the factor so that the week starts on Monday

A8.
```{r}

```


