---
title: "Chapter19"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 19. "Functions" [Chapter 15 hardcopy]
### 19.2.1 Practice

Q1.

* Why is `TRUE` not a parameter to `rescale01()`?
* What would happen if `x` contained a single missing value, and `na.rm` was `FALSE`?

A1.
This refers to a comment in s19.2 R4DS book, (where the `rescale01` function later gets defined), after code that includes `na.rm = TRUE` : "This code only has one input: `df$a`. (If you’re surprised that `TRUE` is not an input, you can explore why in the exercise below.)"

* `TRUE` not a parameter to `rescale01()` because XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX

* if `x` contained a single missing value, and `na.rm` was `FALSE` :
```{r}
# "it’s a good idea to check your function with a few different inputs" -- s19.2 R4DS book :
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)             # na.rm = TRUE
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(1, 2, 3, NA, 5))                # [1] 0.00 0.25 0.50   NA 1.00

# function amended, then retested with the same input as above :
rescale01 <- function(x) {
  rng <- range(x, na.rm = FALSE)            # `na.rm = FALSE`
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(1, 2, 3, NA, 5))                # [1] NA NA NA NA NA
```

"If there were any `NA` values, and `na.rm = FALSE`, then the function would return `NA`" -- [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function]


-----------

Q2.
In the second variant of `rescale01()`, infinite values are left unchanged.
Rewrite `rescale01()` so that `-Inf` is mapped to 0, and `Inf` is mapped to 1

A2.
This is the "second variant of `rescale01()`", from s19.2 R4DS book :
```{r}
x <- c(1:10, Inf)

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(x)
```

Rewriting it [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function] :
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  y <- (x - rng[1]) / (rng[2] - rng[1])            # assigned to `y`
  y[y == -Inf] <- 0                                # `-Inf` is mapped to 0
  y[y == Inf] <- 1                                 # `Inf` is mapped to 1
  y                                                # output
}

rescale01(c(Inf, -Inf, 0:5, NA))                   # [1] 1.0 0.0 0.0 0.2 0.4 0.6 0.8 1.0  NA
```

Note that the regular numbers here (0:5) still get scaled 0:1, even though we have -Inf:Inf


-----------

Q3.
Practice turning the following code snippets into functions. Think about what each function does.
What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?
```{r}
mean(is.na(x))                                       # Q3.1
x / sum(x, na.rm = TRUE)                             # Q3.2
sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)          # Q3.3
```

A3.
A3.1
```{r}
# `is.na(x)` will be True/False (0/1) for each value in vector `x`. The mean will be between 0&1, and indicates the proportion of values that are `NA`
proportion_na <- function(x) {
  mean(is.na(x))
}

# Confirmed by [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function]
```

A3.2
```{r}
# for each value in vector `x` the function will be between 0&1, and they will total to 1 (?)
weights <- function(x) {
  x / sum(x, na.rm = TRUE)
}

# [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function] called the function `weights`, so I copied that name
```

A3.3
```{r}
# 'standard deviation divided by mean' is defined as the "Coefficient of Variation"
coeff_variation <- function(x) {
  sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
}

# See [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function] for use
```


-----------

Q4.
Follow [http://nicercode.github.io/intro/writing-functions.html] to write your own functions to compute the variance and skew of a numeric vector

A4.
I read the page, including "We used the variance function above, but let’s rewrite it" and "Exercise: define a function to compute skew"

See also : [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#when-should-you-write-a-function]


-----------

Q5.
Write `both_na()`, a function that takes two vectors of the same length and returns the number of positions that have an `NA` in both vectors

A5.
XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX


-----------

Q6.
What do the following functions do? Why are they useful even though they are so short?
```{r}
is_directory <- function(x) file.info(x)$isdir        # Q6.1
is_readable <- function(x) file.access(x, 4) == 0     # Q6.2
```


A6.1
XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX

A6.2
XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX


-----------

Q7.
Read the complete lyrics [https://en.wikipedia.org/wiki/Little_Bunny_Foo_Foo] to “Little Bunny Foo Foo”. There’s a lot of duplication in this song. Extend the initial piping example to recreate the complete song, and use functions to reduce the duplication

A7.
XXXXXXXXXXXXXXX XXXXXXXXXXXX XXXXXXXXXXXX


-----------

## 19. "Functions" [Chapter 15 hardcopy]
### 19.3.1 Exercises

Q1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
```

A1.

A1.1
```{r}
?substr         # "Extract or replace substrings in a character vector" -- Help
                # Usage : `substr(x, start, stop)` -- Help
```
We don't know details of the `prefix` (or `string`) object contents, but `f1` appears to be a check of whether the start of `string` matches something from a list of prefixes. T/F value

"The function `f1` returns whether a function has a common prefix ... A better name for `f1` is `has_prefix()`" -- [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#functions-are-for-humans-and-computers]

A1.2
"The function `f2` drops the last element ... A better name for `f2` is `drop_last()`" -- [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#functions-are-for-humans-and-computers]

A1.3
```{r}
?rep            # "rep replicates the values in x. It is a generic function" -- Help
                # "`length.out` : The desired length of the output vector" -- Help
```
"The function `f3` repeats `y` once for each element of `x` ... This is a harder one to name. I would say something like `recycle` (R’s name for this behavior), or `expand`" -- [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#functions-are-for-humans-and-computers]


-----------

Q2. Take a function that you’ve written recently and spend 5 minutes brainstorming a better name for it and its arguments.

A2.

-----------

Q3. Compare and contrast `rnorm()` and `MASS::mvrnorm()`. How could you make them more consistent?

A3.
```{r}
?rnorm           # "'The Normal Distribution' ... eg. `rnorm(n, mean = 0, sd = 1)`" -- Help
                 # "The length of the result is determined by `n` for `rnorm`" -- Help

?MASS::mvrnorm   # "Produces one or more samples from the specified multivariate normal distrib"
                 # `mvrnorm(n = 1, mu, Sigma, tol = 1e-6, empirical = FALSE, EISPACK = FALSE)`" -- Help
```

Based on [https://jrnold.github.io/r4ds-exercise-solutions/functions.html#functions-are-for-humans-and-computers] :
* `rnorm` samples from the univariate normal distribution, while `MASS::mvrnorm` samples from the multivariate normal distribution. The main arguments in `rnorm` are `n`, `mean`, `sd`. But in `MASS::mvrnorm` they are `n`, `mu`, `Sigma`
* To be consistent they should have the same names. And in general, it's better to be consistent with the more-widely used function (ie. `MASS::mvrnorm` should follow the conventions of `rnorm`). However, this is difficult : while `mean` would make sense in the multivariate case, `sd` does not. (Both functions an internally consistent. It wouldn't be good practice to have/mix `mu` & `sd` as arguments, or `mean` & `Sigma`)"


-----------

Q4. Make a case for why `norm_r()`, `norm_d()` etc would be better than `rnorm()`, `dnorm()`. Make a case for the opposite.

A4.


-----------

## 19. "Functions" [Chapter 15 hardcopy]
### 19.4.4 Exercises

Q1. What’s the difference between `if` and `ifelse()`? Carefully read the help and construct three examples that illustrate the key differences.

A1.

-----------

Q2.
Write a greeting function that says “good morning”, “good afternoon”, or “good evening”, depending on the time of day. (Hint: use a time argument that defaults to `lubridate::now()`. That will make it easier to test your function.)

A2.

-----------

Q3.
Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.

A3.

-----------

Q4.
How could you use `cut()` to simplify this set of nested if-else statements?
```{r}
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}
```

How would you change the call to `cut()` if I’d used `<` instead of `<=`? What is the other chief advantage of `cut()` for this problem? (Hint: what happens if you have many values in `temp`?)

A4.


-----------

Q5. What happens if you use `switch()` with numeric values?

A5.

-----------

Q6.
What does this `switch()` call do? What happens if `x` is “e”?
```{r}
switch(x, 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
)
```

Experiment, then carefully read the documentation.

A6.


-----------

## 19. "Functions" [Chapter 15 hardcopy]
### 19.5.5 Exercises

Q1.

* What does `commas(letters, collapse = "-")` do?
* Why?

A1.

-----------

Q2.
It’d be nice if you could supply multiple characters to the pad argument, e.g. `rule("Title", pad = "-+")`

* Why doesn’t this currently work?
* How could you fix it?

A2.

-----------

Q3.

* What does the `trim` argument to `mean()` do?
* When might you use it?

A3.

-----------

Q4.
The default value for the `method` argument to `cor()` is `c("pearson", "kendall", "spearman")`

* What does that mean?
* What value is used by default?

A4.




