---
title: "Chapter10"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
```


## 10. "Tibbles" [Chapter 7 hardcopy]
### 10.5 Exercises
  
Q1. How can you tell if an object is a tibble?
(Hint: try printing `mtcars`, which is a regular data frame)

A1.
```{r}
print(mtcars)
print(as_tibble(mtcars))
```

"Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen. This makes it much easier to work with large data. In addition to its name, each column reports its type, a nice feature borrowed from `str()`" -- s10.3.1 R4DS book

-----------

Q2.
Compare and contrast the following operations on a `data.frame` and equivalent tibble.
What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x                                     # Q2.1
df[, "xyz"]                              # Q2.2
df[, c("abc", "xyz")]                    # Q2.3
```

A2.
```{r}
dftibb <- as_tibble(data.frame(abc = 1, xyz = "a"))
dftibb$x                                 # A2.1
dftibb[, "xyz"]                          # A2.2
dftibb[, c("abc", "xyz")]                # A2.3
```

A2.1
"Compared to a `data.frame`, tibbles are more strict: they never do partial matching, and they will generate a warning if the column you are trying to access does not exist" -- s10.3.2 R4DS book

So the tibble says "Unknown or uninitialised column: 'x'". But the `data.frame` finds `xyz`.
Frustration: This saves keystrokes, but might return (the value of) a 'wrong'/unintended variable

A2.2 & A2.3
"With base R data frames, `[` sometimes returns a data frame, and sometimes returns a vector <if 1 col>. With tibbles, `[` always returns another tibble" -- s10.4 R4DS book

So the tibble returns a tibble (1x1 & 1x2). The `data.frame` returns a vector (because 1 col) & `data.frame` (1x2).
Frustration: Output able to be either vector or `data.frame` means code acting on it has to determine which (else error)

-----------

Q3.
If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?

A3.
You can use the double bracket, like `df[[var]]`.
You can't use the dollar sign, because `df$var` would look for a column named `var`
-- https://jrnold.github.io/r4ds-exercise-solutions/tibbles.html#exercises-13

-----------

Q4.
Practice referring to non-syntactic names in the following data frame by:
  
Q4.1 Extracting the variable called `1`
Q4.2 Plotting a scatterplot of `1` vs `2`
Q4.3 Creating a new column called `3` which is `2` divided by `1`
Q4.4 Renaming the columns to `one`, `two` and `three`

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

A4.
"It’s possible for a tibble to have column names that are not valid R variable names, aka 'non-syntactic' names. For example, they might not start with a letter, or they might contain unusual characters like a space. To refer to these variables, you need to surround them with backticks, `" -- s10.2 R4DS book

A4.1 [Why do those without backticks work? And why doesn't the 2nd one work?] :
```{r}
annoying$`1`           # OK
annoying[[`1`]]        # not ok ...WHY?? ...<Error in `[[.tbl_df`(annoying, `1`) : object '1' not found>
annoying$"1"           # OK ...Why?
annoying[["1"]]        # OK ...Why?
```

A4.2
```{r}
ggplot(annoying, aes(`2`,`1`)) + geom_point()      # Conventional order is `y` vs `x`. So assume `aes(x=`2`, y=`1`)`
```

A4.3
"To use these in a pipe, you’ll need to use the special placeholder ." -- s10.3.2 R4DS book
```{r}
(annoying %>% mutate(`3` = `2`/`1`))    # Above comment does NOT apply here - not subsetting (so no leading ` or $)
```

A4.4 [Renaming the columns to `one`, `two` and `three`] :
```{r}
(annoying %>%
    mutate(`3` = `2`/`1`) %>%                     # Upto here is same as for A4.3
    rename(one = `1`, two = `2`, three = `3`))    # Met `rename()` in s5.4 of R4DS book ..."a variant of `select()`"
```

-----------

Q5. What does `tibble::enframe()` do? When might you use it?

A5.
```{r}
?tibble::enframe()
```
"`enframe()` converts named atomic vectors or lists to two-column data frames" (with names and values) -- Help

"You might use it if you have data stored in a named vector and you want to add it to a data frame and preserve both the name attribute and the actual value" -- http://cfss.uchicago.edu/r4ds_solutions.html
eg.
```{r}
enframe(c(a = 1, b = 2, c = 3))
```

-----------

Q6. What option controls how many additional column names are printed at the footer of a tibble?

A6.
[Note: This answer is relevant if an R script was run. But from an RMarkdown blob, just scroll across for other columns - there is no footer]


"you can explicitly `print()` the data frame and control the number of rows (`n`) and the width of the display. `width = Inf` will display all columns" -- s10.3.1 R4DS book
So, the number of columns displayed will affect how many (others) are printed at the footer
eg.
```{r}
nycflights13::flights %>% print(n = 10, width = 80)    # `width` in characters (not qty of columns)
```


`n_extra` : columns to print abbreviated information for, if the width is too small for the entire tibble. If NULL, the default, it will print information about at most `tibble.max_extra_cols` extra columns
eg.
```{r}
getOption("tibble.max_extra_cols")                     # Tells current value of `tibble.max_extra_cols`
options(tibble.max_extra_cols = 100)                   # Set to 100
```


"You can also control the default print behaviour by setting options: ...
Use `options(tibble.width = Inf)` to always print all columns, regardless of the width of the screen" -- s10.3.1 R4DS book


