---
title: "Chapter13"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(nycflights13)
```


## 13. "Relational Data with dplyr" [Chapter 10 hardcopy]
### 13.2.1 Exercises

Q1.
Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination.
What variables would you need? What tables would you need to combine?

A1.
origin airport ---> destination airport. (This is just the endpoints, but assume it's what 'approximate route' meant)
```{r}
flights      # look at table to see the variables
```

`origin`, `dest` ...[`flights` table]
`lat`, `lon` ...[`airports` table]

Combine `flights` table (`origin`) with `airports` table (`faa`) to get lat&lon of origin airport
Combine `flights` table (`dest`) with `airports` table (`faa`) to get lat&lon of destination airport

-----------

Q2.
I forgot to draw the relationship between `weather` and `airports`.
What is the relationship and how should it appear in the diagram?

A2.
Similarly to Q1, `origin` (in `weather`) could match with `faa` (in `airports`)
(As Q3 says, `weather` only contains data for the NYC airports, not for the destinations)

-----------

Q3.
`weather` only contains information for the origin (NYC) airports.
If it contained weather records for all airports in the USA, what additional relation would it define with `flights`?

A3.
Similarly to Q1, `origin` (in `weather`) could match with `origin` or `dest` (in `flights`) ...`dest` would be the additional part. 
But to know weather for a flight, would also need to match a specific `year`, `month`, `day`, `hour` (and the ease of that may vary depending on frequency/accuracy of readings)

-----------

Q4.
We know that some days of the year are “special”, and fewer people than usual fly on them.

* How might you represent that data as a data frame? What would be the primary keys of that table?
* How would it connect to the existing tables?

A4.
Inference is that these 'special' days would be in a new dataframe/ table. They might be on different dates each year.

* So `year`, `month`, `day` in the dataframe would combine for a primary key of `date`.
* It would connect to `flights` via `year`, `month`, `day`


----------

13. "Relational data" [Chapter 10 hardcopy]
13.3.1 Exercises

Q1.
Add a surrogate key to `flights`

A1.
"If a table lacks a primary key, it’s sometimes useful to add one with `mutate()` and `row_number()`. ... This is called a surrogate key" -- s13.3 R4DS book

```{r}
flights %>%
  mutate(flight_id = row_number())
```

-----------

Q2.
Identify the keys in the following datasets

Q2.1 `Lahman::Batting`
Q2.2 `babynames::babynames`
Q2.3 `nasaweather::atmos`
Q2.4 `fueleconomy::vehicles`
Q2.5 `ggplot2::diamonds`

(You might need to install some packages and read some documentation.)

A2.
In Anaconda, installed the package : **r-lahman**

```{r}
library(Lahman)
# library(babynames)               # No. Package wasn't in the Anaconda distribution
# library(nasaweather)             # No. Package wasn't in the Anaconda distribution
# library(fueleconomy)             # No. Package wasn't in the Anaconda distribution
```

A2.1
```{r}
Lahman::Batting

Batting %>%
  count(playerID, yearID, stint) %>%
  filter(n > 1)                    # 13876 rows if trying `playerID` as the key. 7157 with `(playerID, yearID)`
```
The combined primary key is `(playerID, yearID, stint)`

A2.2 -     # Package wasn't in the Anaconda distribution
A2.3 -     # Package wasn't in the Anaconda distribution
A2.4 -     # Package wasn't in the Anaconda distribution

A2.5
```{r}
ggplot2::diamonds
```
There is no primary key here. From [https://jrnold.github.io/r4ds-exercise-solutions/relational-data.html#keys] we learn that "Using all variables in the data frame, the number of distinct rows is less than the total number of rows, meaning no combination of variables uniquely identifies the observations". (ie. Some rows are exactly the same as other rows.) We could create a surrogate key with `row_number()`, as in Q1

-----------

Q3.
Q3.1 Draw a diagram illustrating the connections between the `Batting`, `Master`, and `Salaries` tables in the **Lahman** package.
Q3.2 Draw another diagram that shows the relationship between `Master`, `Managers`, `AwardsManagers`.
Q3.3 How would you characterise the relationship between the `Batting`, `Pitching`, and `Fielding` tables?

A3.
A3.1
```{r}
Batting                                 # Primary key is `(playerID, yearID, stint)` ...[from A2.1]
Master                                  # Primary key is `playerID`
Salaries

Salaries %>%
  count(yearID, teamID, playerID) %>%   # Primary key is `(yearID, teamID, playerID)`
  filter (n > 1)
```

Master$playerID matches with Batting$playerID ...(1:Many)
Master$playerID matches with Salaries$playerID ...(1:Many)

A3.2 [reminder : "relationship between `Master`, `Managers`, `AwardsManagers`"] :
```{r}
?Lahman

Master                                  # Primary key is `playerID` ...[from A3.1]
Managers                                # Primary key is `(yearID, teamID, inseason)`
AwardsManagers                          # Primary key is `(playerID, yearID, awardID)`


Managers %>%
  count(yearID, teamID, inseason) %>%
  filter(n >1)
# `inseason` : "Managerial order. Zero if the individual managed the team the entire year. Otherwise denotes where the manager appeared in the managerial order (1 for first manager, 2 for second, etc.)" -- Help. ...`inseason` is needed in the primary key, because a `playerID` might have managed the same team more than once in a year (eg. 1st & 3rd), but `playerID` is then NOT needed

AwardsManagers %>%
  count(playerID, yearID, awardID) %>%
  filter (n > 1)
```

Master$playerID matches with Managers$playerID ...(1:Many)
Master$playerID matches with AwardsManagers$playerID ...(1:Many)

A3.3 [reminder : "relationship between the `Batting`, `Pitching`, and `Fielding` tables?"] :
"[These] tables all have a primary key consisting of the `playerID`, `yearID`, and `stint` variables. They all have a 1-1 relationship to each other" -- https://jrnold.github.io/r4ds-exercise-solutions/relational-data.html#keys


----------

13. "Relational data" [Chapter 10 hardcopy]
13.4.6 Exercises

Q1.
Compute the average delay by destination, then join on the `airports` data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
```

(Don’t worry if you don’t understand what `semi_join()` does — you’ll learn about it next.)
You might want to use the size or colour of the points to display the average delay for each airport

A1.
[Didn't work when I named the variable `avg-delay` instead of `avg_delay`. '-' isn't a valid character. It could be considered as a minus sign between variables named `avg` and `delay` ...though probably `avg` is restricted]
```{r}
flights %>%
  group_by(dest) %>%                                          # `group_by()` in order to then take the group average
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%    # Choose to exclude `NA`s from the average
  left_join(airports, by = c(dest = "faa"))                   # Join using flight destination, not flight origin
```

This next part is same as in the question, but colours each airport (by `avg_delay`) :
```{r}
ggplot(aes(lon, lat, colour = avg_delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
```

"The left join should be your default join: use it unless you have a strong reason to prefer one of the others" -- s13.4.3 R4DS book
"Warning message: Removed 4 rows containing missing values (geom_point)". So did it plot as if I'd chosen `inner_join`?

-----------

Q2.
Add the location of the origin _and_ destination (i.e. the `lat` and `lon`) to `flights`

A2.
Answer combines 2 pieces of example code at the end of s13.4.5. Looks as if it worked, but is it the right way?
...(ie. one after the other in the 'and then' pipe, rather than some sort of parallel? ...if that is even possible)
```{r}
flights %>%
  left_join(airports, by = c(origin = "faa")) %>%  # Looks like the join used in A1
  left_join(airports, by = c(dest = "faa")) %>%    # `airports` table doesn't distinguish between origin&destination
View()
```

There are name duplications from the 2 joins, but each has a suffix (assume .x is `origin` and .y is `dest`)
...re. "disambiguated in the output with a suffix" (s13.4.5 R4DS book, where 2 tables each had same name for different variable)

-----------

Q3.
Is there a relationship between the age of a plane and its delays?

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q4. What weather conditions make it more likely to see a delay?
A4.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q5.
What happened on June 13 2013?
Display the spatial pattern of delays, and then use Google to cross-reference with the weather

A5.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX


----------

13. "Relational data" [Chapter 10 hardcopy]
13.5.1 Exercises

Q1.
What does it mean for a flight to have a missing `tailnum`?
What do the tail numbers that don’t have a matching record in planes have in common?
(Hint: one variable explains ~90% of the problems.)

A1.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q2.
Filter `flights` to only show flights with planes that have flown at least 100 flights

A2.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q3.
Combine `fueleconomy::vehicles` and `fueleconomy::common` to find only the records for the most common models

A3.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q4.
Find the 48 hours (over the course of the whole year) that have the worst delays.
Cross-reference it with the weather data. Can you see any patterns?

A4.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q5.
What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you?
What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?

A5.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

-----------

Q6.
You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above

A6.
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
