---
title: "Chapter15"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(forcats)
```


## 15. "Factors" [Chapter 12 hardcopy]
### 15.3.1 Exercises

Q1.
Explore the distribution of `rincome` (reported income).
What makes the default bar chart hard to understand?
How could you improve the plot?

A1.
```{r}
?gss_cat
# "It’s a sample of data from the General Social Survey, which is a long-running US survey" -- s15.3 R4DS book

forcats::gss_cat %>%
  ggplot(aes(x = rincome)) +          # "Explore the distribution of `rincome`" per Q1.
  geom_bar()
```

Here, "hard to understand" refers to the x-axis labels being squashed/overlapping/unreadable. Usual way to improve this (unless bars have to be vertical) is to make the bars horizontal :
```{r}
forcats::gss_cat %>%
  ggplot(aes(rincome)) +
  geom_bar() +
  coord_flip()                        # "switches the x and y axes" -- s3.9 R4DS book
```

Other possible improvements to consider :

* descending values (ie. reading down an ordered list of `count`) ...but "`rincome` already has a principled order that we shouldn’t mess with. Reserve `fct_reorder()` for factors whose levels are arbitrarily ordered" -- s15.5 of R4DS book
* consistent bin size ...(but we can't affect that)
* ...and maybe grouping some 'no answer' categories, if no values can be got from them

-----------

Q2.
What is the most common `relig` in this survey?
What’s the most common `partyid`?

A2.
```{r}
forcats::gss_cat %>%
  count(relig) %>%
  arrange(desc(n))                             # s5.3 R4DS book
# Most common `relig` is Protestant

forcats::gss_cat %>%
  count(partyid) %>%                           # `partyid` instead of `relig`
  arrange(desc(n))
# Most common `partyid` is Independent
```

-----------

Q3.
Which `relig` does `denom` (denomination) apply to?
How can you find out with a table?
How can you find out with a visualisation?

A3.
```{r}
forcats::gss_cat$denom          # Values for first 1000 rows. Answer is Protestant?
levels(forcats::gss_cat$denom)  # 30 different values. More confident in answer of Protestant

forcats::gss_cat %>%
  ggplot(aes(relig, denom)) +   # With `relig` on other axis, can see the mapping. Protestant
  geom_point() +
  coord_flip()            # Incl/excl this, to read both axis labels. (Q implied only 1 answer)

# Answer from [https://jrnold.github.io/r4ds-exercise-solutions/factors.html#general-social-survey] displays quantities as size, but its most useful feature is rotation of labels on the x-axis (so not necessary to keep 'flipping') :
gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))        # rotates x-axis labels by 90dg
```


-----------

## 15. "Factors" [Chapter 12 hardcopy]
### 15.4.1 Exercises
  
Q1. There are some suspiciously high numbers in `tvhours`. Is the mean a good summary?

A1.
```{r}
forcats::gss_cat %>%
  ggplot(aes(x = tvhours)) +
  geom_histogram(binwidth = 1)
# "Warning message: Removed 10146 rows containing non-finite values (stat_bin)"

# Answer from [https://jrnold.github.io/r4ds-exercise-solutions/factors.html#modifying-factor-order] :
gss_cat %>%
  filter(!is.na(tvhours)) %>%       # extra line means no 'Warning' received (but same chart)
  ggplot(aes(x = tvhours)) +
  geom_histogram(binwidth = 1)
```

* The distribution has a long tail to the right, but most results are <6 hrs, and nothing >24 hrs
* Median might be a 'better' summary, but also 'it depends' (on the reason for wanting a summary)

-----------

Q2. For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled.

A2.
"`rincome` already has a principled order that we shouldn’t mess with. Reserve `fct_reorder()` for factors whose levels are arbitrarily ordered" -- s15.4 R4DS book
```{r}
?gss_cat
# variables : year, marital, age, race, rincome, partyid, relig, denom, tvhours
# categorical variables : marital, race, rincome, partyid, relig, denom

summary(gss_cat)
# factors : marital, race, rincome, partyid, relig, denom
# arbitrary order : race, relig, denom
# principled order : marital?, rincome, partyid?
```

Answer from [https://jrnold.github.io/r4ds-exercise-solutions/factors.html#modifying-factor-order] goes into more detail, with text & plots, and says `race` *does* have a principled order (based on count of observations)

-----------

Q3. Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

A3.
The question is referring to the last `geom_point()` chart in s15.4
```{r}
ggplot(rincome_summary, aes(age, fct_relevel(rincome, "Not applicable"))) +
  geom_point()
# (For reference. We haven't defined `rincome_summary` in this document, so it will error)
```

* Order of the levels in the `rincome()` factor can be seen with `levels(gss_cat$rincome)` or `gss_cat %>% count(rincome)`, etc
* R assigns integers to each level, based on the order. So "1" to the 1st level, "2" to the 2nd level, etc
* When plotting `rincome` on the y-axis, the level nearest the '0' will be that with integer "1", etc

```{r}
?fct_relevel
```

Using `fct_relevel` to move "Not applicable" to 1st level gives it integer "1" (and "move it the bottom of the plot")


-----------

## 15. "Factors" [Chapter 12 hardcopy]
### 15.5.1 Exercises

Q1. How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

A1.
```{r}
forcats::gss_cat %>% 
  count(year, partyid) %>%
  # ...Without this, `year` & `partyid` not recognised in `ggplot()`. Not sure why
  group_by(year) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(year, proportion, colour = partyid)) +         # (10 coloured lines for `partyid`)
  geom_line()
```

Answer from [https://jrnold.github.io/r4ds-exercise-solutions/factors.html#modifying-factor-levels] also :

* Grouped the many `partyid` values into a succinct Democrat/ Republican/ Independent/ Other ...with `fct_collapse()`
* And reordered the Legend ...with `fct_reorder2()`
* And showed each year's datapoints ...with `geom_point()`
* And renamed the Legend from `partyid` to "Party ID." ...with `labs()`
```{r}
gss_cat %>% 
  mutate(partyid = 
           fct_collapse(partyid,
                        other = c("No answer", "Don't know", "Other party"),
                        rep = c("Strong republican", "Not str republican"),
                        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
                        dem = c("Not str democrat", "Strong democrat"))) %>%
  count(year, partyid)  %>%
  group_by(year) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = year, y = p,
             colour = fct_reorder2(partyid, year, p))) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID.")
```


-----------

Q2.
How could you collapse `rincome` into a small set of categories?

A2.
```{r}
forcats::gss_cat$rincome                # Values for first 1000 rows
levels(forcats::gss_cat$rincome)        # 30 different values in total

# Group 'non-responses' together into one category :
forcats::gss_cat %>%
  mutate(rincome = fct_collapse(
                    rincome,
                    `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable")
                    )) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() + 
  coord_flip()
```


Answer from [https://jrnold.github.io/r4ds-exercise-solutions/factors.html#modifying-factor-levels] also created 'consistent bin size' ...(re. 15.3.1 Exercises Q1) ...with `fct_collapse()` after using `str_c()` :
```{r}
library("stringr")
gss_cat %>%
  mutate(rincome = 
           fct_collapse(
             rincome,
             `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
             `Lt $5000` = c("Lt $1000", str_c("$", c("1000", "3000", "4000"),
                                              " to ", c("2999", "3999", "4999"))),
             `$5000 to 10000` = str_c("$", c("5000", "6000", "7000", "8000"),
                                      " to ", c("5999", "6999", "7999", "9999"))
           )) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() + 
  coord_flip()
```


(re. Answer above.) I don't follow how piecing together a heading with `str_c()` will 'grab' the records it relates to. ...So, this answer [below] also works to create 'consistent bin size', and makes sense to me :
```{r}
library("stringr")
gss_cat %>%
  mutate(rincome = 
           fct_collapse(
             rincome,
             `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
             `Lt $5000` = c("Lt $1000", "$1000 to 2999", "$3000 to 3999", "$4000 to 4999"),
             `$5000 to 10000` = c("$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999")
           )) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() + 
  coord_flip()
```



