---
title: "Chapter28"
output:
  github_document: default
  html_notebook: default
---

```{r}
library("tidyverse")
library("modelr")
library("lubridate")
```


## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.2.1 Exercises

Q1.
Create one plot on the fuel economy data with customised `title`, `subtitle`, `caption`, `x`, `y`, and `colour` labels

A1.
Combining the examples from s28.2 R4DS book (which corrects/removes `paste()` code that is in the hardcopy) :

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```


-----------

Q2.
The `geom_smooth()` is somewhat misleading because the `hwy` for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model

A2.
Starting from the code in A2 and making 2 ADDitions, then revisualising :

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +      # ADD: `colour = class` to get multiple lines
  geom_point(aes(colour = class)) +
  geom_smooth(method = "lm", se = FALSE) +          # ADD: `method = "lm"` to get straight lines
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

Based on [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#label] (but that answer had relabelled x&y axes the wrong way around) :

```{r}
mod <- lm(hwy ~ class, data = mpg)             # model based on fuel economy for Class (Car type)
mpg %>%
  add_residuals(mod) %>%
  ggplot(aes(displ, resid)) +                  # `resid` is default column name for residuals
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Highway MPG for Cars After Subtracting Mean MPG of their Class",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy, relative to Class (mpg)"
  )
```

[Not sure what this is intended to show. I assume it's that after removing effects of all the coloured linear models, we're left with this one that's fairly 'flat' (and so there really is a linear relationship for each `Car type`)? Otherwise, I assume it's that the coloured models shouldn't be straight lines?]


-----------

Q3.
Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

A3.


-----------

## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.3.1 Exercises

Q1. Use `geom_text()` with infinite positions to place text at the four corners of the plot

A1.
```{r}
label_bottomleft <- tibble(displ = -Inf, hwy = -Inf, label = "Bottom left")
label_topleft <- tibble(displ = -Inf, hwy = Inf, label = "Top left")
label_topright <- tibble(displ = Inf, hwy = Inf, label = "Top right")
label_bottomright <- tibble(displ = Inf, hwy = -Inf, label = "Bottom right")

# (Based on plot used near end of s28.3.) This places the text ok, but it's mostly off-screen :
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label_bottomleft) +
  geom_text(aes(label = label), data = label_topleft) +
  geom_text(aes(label = label), data = label_topright) +
  geom_text(aes(label = label), data = label_bottomright)

# This is longer, but justifies the text, so that it's all readable :
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label_bottomleft, hjust = "left", vjust = "bottom") +
  geom_text(aes(label = label), data = label_topleft, hjust = "left", vjust = "top") +
  geom_text(aes(label = label), data = label_topright, hjust = "right", vjust = "top") +
  geom_text(aes(label = label), data = label_bottomright, hjust = "right", vjust = "bottom")
```

[https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#annotations] shows a way to simplify this, by creating a table in a `tribble()`, and then using "a single `geom_text` call instead of four different `geom_text` calls with four different `data` arguments, and four different values of `hjust` and `vjust` arguments"


-----------

Q2.
Read the documentation for `annotate()`. How can you use it to add a text label to a plot without having to create a tibble?

A2.
```{r}
?ggplot2::annotate

# "This function adds geoms to a plot, but unlike typical a geom function, the properties of the geoms are not mapped from variables of a data frame, but are instead passed in as vectors. This is useful for adding small annotations (such as text labels) or if you have your data in vectors, and for some reason don't want to put them in a data frame" -- Help
```

Based on [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#annotations] : "With `annotate` you use what would be aesthetic mappings directly as arguments" :
```{r}
ggplot(mpg, aes(displ, hwy)) +                           # same as before
  geom_point() +                                         # same as before
  annotate("text", x = Inf, y = Inf,                     # instead of `geom_text()`
           label = "Top right again", vjust = "top", hjust = "right"
           )
```


-----------

Q3.
How do labels with `geom_text()` interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

A3.
Based on [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#annotations] (which also shows the plots) :

* If the faceting variable isn't in the `label` data frame, the same text is put in all facets
* To 'add a label to a single facet', add a column to the `label` data frame for the faceting variable, along with the value of the single facet. [eg. `class = "2seater"`]
* To put 'a different label in each facet', use the faceting variable in the `label` dataframe.   [eg. `class = unique(mpg$class)`]

-----------

Q4. What arguments to `geom_label()` control the appearance of the background box?

A4.
```{r}
?geom_label

# "`geom_text` adds text directly to the plot. `geom_label` draws a rectangle behind the text, making it easier to read" -- Help
```

* `label.padding`: Amount of padding around label. Defaults to 0.25 lines
* `label.r`: Radius of rounded corners. Defaults to 0.15 lines
* `label.size`: Size of label border, in mm

"The `fill` aesthetic controls the background colour of the label"


-----------

Q5.
What are the four arguments to `arrow()`? How do they work? Create a series of plots that demonstrate the most important options

A5.
```{r}
?grid::arrow

# "Produces a description of what arrows to add to a line. The result can be passed to a function that draws a line, eg. `grid.lines`" -- Help
```

* `angle` : of the arrow head in degrees (smaller numbers produce narrower, pointier arrows)
* `length` : A unit specifying the length of the arrow head (from tip to base)
* `ends`: "last", "first", or "both", indicating which ends of the line to draw arrow heads
* `type`: "open" or "closed" indicating whether the arrow head should be a closed triangle


-----------

## 28. "Graphics for communication" [Chapter 22 hardcopy]
### 28.4.4 Exercises

Q1.
Why doesn’t the following code override the default scale? :
```{r}
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_colour_gradient(low = "white", high = "red") +             # _COLOUR_
  coord_fixed()
```

A1.
From [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#scales] :
"It does not override the default scale because the colors in `geom_hex` are set by the `fill` aesthetic, not the `color` aesthetic"

```{r}
# df <- tibble(
#   x = rnorm(10000),
#   y = rnorm(10000)
# )

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +               # _FILL_
  coord_fixed()
```


-----------

Q2. What is the first argument to every scale? How does it compare to `labs()`?

A2.
```{r}
?scale_x_continuous
# [The first argument is `name`,] "the name of the scale. Used as axis or legend title. If NULL, the default, the name of the scale is taken from the first mapping used for that aesthetic" -- Help

?labs
# [The arguments are] "A list of new name-value pairs. The name should either be an aesthetic, or one of 'title', 'subtitle', or 'caption' -- Help
```

Based on [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#scales] :
"The first argument to every scale is the name for the scale. It is equivalent to using `labs()`"



-----------

Q3.
Change the display of the presidential terms by:

1. Combining the two variants shown above
2. Improving the display of the y axis
3. Labelling each term with the name of the president
4. Adding informative plot labels
5. Placing breaks every 4 years (this is trickier than it seems!)

A3.
The question is referring to this code from s28.4.3 R4DS book. (Don't know what 'the two variants' refers to) :
```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

From [https://jrnold.github.io/r4ds-exercise-solutions/graphics-for-communication.html#scales] :
```{r}
years <- lubridate::make_date(seq(year(min(presidential$start)),
             year(max(presidential$end)),
             by = 4), 1, 1)                                       # 'every 4 years'

presidential %>%
  mutate(id = 33 + row_number(),
         name_id = stringr::str_c(name, " (", id, ")"),           # added
         name_id = factor(name_id, levels = name_id)) %>%         # added
  ggplot(aes(start, name_id, colour = party)) +                   # `name_id` instead of `id` [3]
    geom_point() +
    geom_segment(aes(xend = end, yend = name_id)) +               # `name_id` instead of `id` [3]
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue")) +
    scale_y_discrete(NULL) +                                      # added, to remove title [2]
    scale_x_date(NULL, breaks = years, date_labels = "'%y") +     # added breaks [5], labels [4]
    theme(panel.grid.minor = element_blank())                     # added
```


-----------

Q4.
Use `override.aes` to make the legend on the following plot easier to see

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20)
```

A4.
Colours in the `cut` legend all appear to be v v faint pastel (for Fair, Good, Very Good, Premium, Ideal), because the `alpha` is so small (5%)

From s28.4.2 R4DS book :
"The following example shows two important settings: controlling the number of rows the legend uses with `nrow`, and overriding one of the aesthetics to make the points bigger. This is particularly useful if you have used a low alpha to display many points on a plot" :
`guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))`


We can use `override.aes` to make the points solid in the legend, without changing them in the plot :
```{r}
ggplot(diamonds, aes(carat, price)) +                                  # no change
  geom_point(aes(colour = cut), alpha = 1/20)  +                       # no change
  guides(colour = guide_legend(override.aes = list(alpha = 1)))        # `alpha = 1`
```




